---
title: "ipRGC bulk RNAseq -- All Samples -- This time with feeling (transcript-level abundances)!"
author: "Theresa Alexander, modified by atb"
date: "4/15/2023, modified 202603"
output:
  html_document:
    code_download: true
    code_folding: hide
    fig_caption: true
    fig_height: 7
    fig_width: 7
    highlight: tango
    keep_md: false
    mode: selfcontained
    number_sections: true
    self_contained: true
    theme: readable
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
      toc_depth: 5
      rmdformats::readthedown:
        code_download: true
        code_folding: show
        df_print: paged
        fig_caption: true
        fig_height: 7
        fig_width: 7
        highlight: tango
        width: 300
        keep_md: false
        mode: selfcontained
        toc_float: true
        BiocStyle::html_document:
          code_download: true
          code_folding: show
          fig_caption: true
          fig_height: 7
          fig_width: 7
          highlight: tango
          keep_md: false
          mode: selfcontained
          toc_float: true
---
<style type="text/css">
body, td {
  font-size: 16px;
}
code.r{
  font-size: 16px;
}
pre {
  font-size: 16px
}
</style>

# Introduction

This document is intended to repeat with minimal modification the work
performed by Theresa when analyzing the RNASeq IPRGC data.  This is
being performed in an apptainer (ugh I hate that name) container and
so I will likely remove some of the libraries invoked in it.

Not all of these are available in the container, and I do not think I have the version
of Theresa's helper_functions.R which was used for this.  I did find a
version of it that appears to work without any problems, though.

At the end of this document, I am going to try to re-implement it
with some minor changes to see if I can more easily reconcile the
results to the tables provided by Colenso.

```{r, message=FALSE, eval=FALSE}
library("hpgltools")
library("ggplot2")
library("patchwork")
library("plotly")
library("RColorBrewer")
library("DT")
library("dplyr")
library("openxlsx")
library("ComplexHeatmap")
library("circlize")
library("gprofiler2")
source("helper_functions.R")
library("gplots")
library("ggrepel")
```


```{r}
#setwd("~/scratch_02/reesyxan/mmusculus_iprgc_2019")
```

# M. musculus

## Annotations

I am using mm38_100.

```{r}
mm_annot <- load_biomart_annotations(species = "mmusculus")
mm_annot <- mm_annot[["annotation"]]
mm_annot[["txid"]] <- paste0(mm_annot[["ensembl_transcript_id"]], ".", mm_annot[["version"]])
rownames(mm_annot) <- make.names(mm_annot[["ensembl_gene_id"]], unique=TRUE)

mm_tx_annot <- mm_annot[["annotation"]]
## yeah, the ID tag in my gff file has transcript: as a prefix, so the featureCounts outputs
## includes it...  I will fix that inpost.
rownames(mm_tx_annot) <- paste0("transcript:", rownames(mm_tx_annot))

tx_gene_map <- mm_annot[, c("txid", "ensembl_gene_id")]
```


So, I now have a table of mouse annotations.

## Hisat2 expressionset

```{r}
mm38_hisat <- create_se("sample_sheets/samplesheet_bulkrnaseq_Opn4.xlsx",
                        gene_info = mm_tx_annot, file_column = "salmon_count_table")
```

# Color Set Up

```{r}
color_condition <- mm38_hisat$initial_metadata$condition
color_condition <- factor(color_condition, levels = c("Retina_het", "Retina_ko",
                                                      "SCN_het", "SCN_ko", "dLGN_het", "dLGN_ko"))

color_values <- c( "#00BFC4",  "#9AD2D3", "#F8766D", "#FBADA7", "#7CAE00", "#D3EE8F")
col_fun = colorRamp2(c(0, 3, 6), c("blue", "pink", "red"))
```

# Mapped Counts Per Sample Plot

```{r}
library(forcats)
mm38_hisat <- set_conditions(mm38_hisat, fact = "batch")
libsize_df <- as.data.frame(plot_libsize(mm38_hisat)$table)
rownames(libsize_df) <- mm38_hisat$design$sampleid
libsize_df$samplenum <- paste0(mm38_hisat$design[rownames(libsize_df), "samplenumb"], "_",
                               mm38_hisat$design[rownames(libsize_df), "batch"])
libsize_df$sample_id <-  rownames(mm38_hisat$design)
libsize_df <- libsize_df[order(libsize_df$sum),]

ggplot(libsize_df, aes(x = fct_rev(fct_infreq(samplenum, sum)), y = sum, fill = condition)) +
  geom_bar(stat = "identity", color = "black") +
  theme_classic() +
  #theme(axis.text.x= element_text(fct_rev(fct_infreq(libsize_df$samplenum, libsize_df$sum))), #remove x axis labels
  #    axis.ticks.x=element_blank()) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_fill_manual(values = c("#60449b", "#009f6f", "orange")) +
  xlab("Sample") +
  ylab("Sum of Mapped Gene Counts") +
  labs(fill = "")

plot_libsize(mm38_hisat)
```

# Normalize Expression

First we will filter out the genes which have low counts across all
samples. Then I will normalize both without and with SVA correction.

```{r}
mm38_filt <- normalize(mm38_hisat, filter = TRUE)
des <- colData(mm38_filt)
new_factor <- paste0(des[["location"]], "_", des[["genotype"]])
colData(mm38_hisat)[["location_genotype"]] <- new_factor

mm38_norm <- normalize(mm38_hisat, filter = TRUE, convert = "cpm",
                       norm = "quant", transform="log2")

mm38_norm <- set_conditions(mm38_norm, fact = "location_genotype")
```

# Melanopsin Expression

Here is a plot of the Opn4 expression across all retina samples split
by batch. The 4 quadrants in the plot are <br>

het:1 -- retina het samples in batch 1 <br>
het:2 -- retina het samples in batch 2 <br>
ko:1 -- retina ko samples in batch 1 <br>
ko:2 -- retina ko samples in batch 2 <br>

This is a Tx comparison; I need to either drop this or hunt down the set of
transcript IDs (which I did)

```{r, eval=FALSE}
library(RColorBrewer)
colors <- brewer.pal(name = "Spectral", n = 2)
mm38_cpm <- normalize(mm38_filt, convert = "cpm")
df <- data.frame(exprs = assay(mm38_cpm)["ENSMUSG00000021799",],
                 genotype = as.factor(des[["genotype"]]),
                 location = as.factor(des[["location"]]),
                 sampleid = des[["sampleid"]],
                 samplenum = des[["samplenumb"]],
                 batch = des[["batch"]])

## Set postscript arguments
setEPS()
postscript("bulkRNAseq_Opn4Expression_Retina.eps")
cpm_df <- df |>
  filter(location == "Retina" | location == "retina")

## This is where I most miss ggstatsplot
## I am going to sit and write my own version of it.
ggplot(cpm_df, aes(x = samplenum, y = exprs, fill = genotype, group = genotype)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8)) +
  xlab("Sample") +
  ylab("Opn4 Counts Per Million (CPMs)") +
  #scale_fill_manual(values = c("#FFC107", "#004D40", "orange"),
  #                  aesthetics = "fill") +
  ggtitle("Retina Melanopsin Expression") +
  facet_wrap(~genotype:batch, scales = "free") +
  ylim(c(0, 25)) +
  theme_classic()
dev.off()
```

The WT samples in batch 3 have about double the Opn4 expression as the
het samples in batch 3 (and batch 2).

# PCA

```{r pca, message = FALSE}
pca_norm <- plot_pca(mm38_norm, plot_labels = FALSE)
pca_norm[["plot"]]
## rownames(pca_norm$result$v)
labels <- des[["sampleid"]]
## labels <- paste0(mm38_norm$design$sampleid, "_", mm38_norm$design$samplenumb)
all_df <-  cbind(as.data.frame(pca_norm$result$v), labels)

shape_genotype <- mm38_norm$design$genotype
color_condition <- mm38_norm$condition

## filter(sample_id %in% c("S3_Het_Ret_S1", "S4_Het_Ret_S4", "S5_Het_Ret_S16", "S1_Het_Ret_S1", "S2_Het_Ret_S4", "S3_Het_Ret_S7", "S4_Het_Ret_S10", "S5_Het_Ret_S13")) |>

pca_norm_plot <-  ggplot(all_df, aes(x = PC1, y = PC2, label = labels)) +
  geom_point(aes(color = color_condition, shape = shape_genotype),
             size = 5) +
  #scale_color_manual(values = color_values,
  #                  aesthetics = "color") +
  theme_classic() +
  labs(color="Cell Compartment/Genotype")   +
  ggtitle("PCA with No Batch Correction")  +
  labs(x = paste0("PC1: ",  pca_norm$prop_var[1], "% variance"),
       y =  paste0("PC2: ",  pca_norm$prop_var[2], "% variance"))

#setEPS()                                             # Set postscript arguments
#postscript("bulkRNAseq_PCA_allsamples_labeledS5.eps")                           # Start graphics device driver
all_df[["labels"]] <- rownames(all_df)
pca_norm_plot + ggtitle("PCA Normalized Counts") +
geom_label_repel(data=filter(all_df,
                             labels %in% labels), #c('s5_het_dlgn', 's5_het_ret', 's5_het_scn')),
    nudge_x = 0.15,
    nudge_y = 0.15)
#dev.off()
```

## Look at the first batch of sequencing

```{r}
#batch 1 plot
all_df[rownames(all_df) %in% tolower(rownames(des[des[["batch"]] == "1",])),] |>
  ggplot(aes(x = PC1, y = PC2)) +
  geom_point(aes(color = color_condition[which(rownames(all_df) %in% tolower(rownames(des[des$batch == "1",])))],
                 shape = shape_genotype[which(rownames(all_df) %in% tolower(rownames(des[des$batch == "1",])))]),
             size = 5) +
  scale_color_manual(values = color_values,
                     aesthetics = "color") +
  theme_classic() +
  labs(color="Cell Compartment/Genotype", shape = "Genotype")   +
  ggtitle("PCA - Batch 1 RNAseq")  +
  labs(x = paste0("PC1: ",  pca_norm$prop_var[1], "% variance"),
       y =  paste0("PC2: ",  pca_norm$prop_var[2], "% variance"))
```

I think there is an easier way to do the above

```{r}
b1 <- subset_se(mm38_norm, subset = "batch == '1'")
plot_pca(b1)

b1_b2 <- subset_se(mm38_norm, subset = "batch == '1'|batch == '2'")
plot_pca(b1_b2)
```

```{r}
#batch 1&2 plot
pca_batch1and2 <- all_df |>
  ggplot(aes(x = PC1, y = PC2)) +
  geom_point(aes(color = color_condition,
                 shape = shape_genotype),
             size = 5) +
  scale_color_manual(values = color_values,
                     aesthetics = "color") +
  theme_classic() +
  labs(color="Cell Compartment/Genotype", shape = "Genotype")   +
  ggtitle("PCA - Batch 1&2 RNAseq")  +
  labs(x = paste0("PC1: ",  pca_norm$prop_var[1], "% variance"),
       y =  paste0("PC2: ",  pca_norm$prop_var[2], "% variance"))


#setEPS()
#postscript(file = "rnaseq_pca-batch1and2.eps")
#pca_batch1and2
#dev.off()
```

In this pca plot, we see the retina samples cluster on the right in
the PC1 direction and the brain samples cluster on the left in the PC
1 direction.

As we saw before, there is disparity between some of the hets. As it
turns out, all of the new Het samples cluster with the KO samples as
our one weirdo did before. So now it seems that it is quite easy to
get some genetic drift that leads to these mice's expression to behave
in weird ways.

As we did before, let's look at the Opn4 expression across all the
retina samples to see if the Het samples that behave like KOs have
lower Opn4 expression.

# A closer look at the similar KOs and Hets

Let's exclude the 2 Hets that are the most dissimilar from the bunch
and look at a PCA plot of the rest of the samples.

```{r}
## The following subset is using an older sample ID which I am unable to find anywhere.
## As a result, I rewrote that portion of Theresa's pca plot and labeled everything
## according to the extant naming scheme, I will now try and figure out which
## samples were removed.
## mm38_subset <- subset_se(
##   mm38_filt,
## subset = "sampleid != 'S3_Het_dLGN'&sampleid != 'S3_Het_Ret'&sampleid != 'S3_Het_SCN'&sampleid != 'S4_Het_dLGN'&sampleid != 'S4_Het_Ret'&sampleid != 'S4_Het_SCN'")
## S3_Het_dLGN_S2 is an outlier marked by Theresa
## S3_Het_Ret_S1 is an outlier
## S3_Het_SCN_S3 is an outlier
## S4_Het_dLGN_S11 is somewhat an outlier, I might be inclined to leave it.
## S4_Het_Ret_S4 is a huge outlier
## S4_Het_SCN_S6 is an outlier

## If I were choosing likely samples, I think I would go for:
##S4_Het_Ret_S4, S3_Het_Ret_S1, S3_Het_SCN_s3, S3_Het_dLGN_S2, S4_Het_dLGN_S5, S4_Het_SCN_S6

## Either way, I am adding a couple columns to the sample sheet with these calls for included samples.

mm38_subset <- subset_se(mm38_filt, subset = "included=='yes'")

mm38_subset <- normalize(mm38_subset, filter=TRUE, convert="cpm",
                         norm = "quant", transform="log2")

mm38_subset <- set_conditions(mm38_subset, fact = "location_time")

pca_norm_subset <- plot_pca(mm38_subset, plot_labels = FALSE)
pca_norm_subset

labels <- rownames(pca_norm_subset$result$v)

all_df <-  cbind(as.data.frame(pca_norm_subset$result$v), labels)
shape_genotype <- mm38_subset$design$genotype
color_condition <- pca_norm_subset$table$condition

pca_norm_plot <-  ggplot(pca_norm_subset$table, aes(x = PC1, y = PC2)) +
  geom_point(aes(color = condition),
             size = 5) +
  scale_color_manual(values = color_values,
                     aesthetics = "color") +
  theme_classic() +
  labs(color="Cell Compartment/Genotype")   +
  ggtitle("PCA with No Batch Correction")  +
  labs(x = paste0("PC1: ",  pca_norm_subset$prop_var[1], "% variance"),
       y =  paste0("PC2: ",  pca_norm_subset$prop_var[2], "% variance"))

pca_norm_plot <-  ggplot(all_df, aes(x = PC1, y = PC2, label = labels)) +
  geom_point(aes(color = color_condition, shape = shape_genotype),
             size = 5) +
  #scale_color_manual(values = color_values,
  #                  aesthetics = "color") +
  theme_classic() +
  labs(color="Cell Compartment/Genotype")   +
  ggtitle("PCA with No Batch Correction")  +
  labs(x = paste0("PC1: ",  pca_norm_subset$prop_var[1], "% variance"),
       y =  paste0("PC2: ",  pca_norm_subset$prop_var[2], "% variance"))

#setEPS()                                             # Set postscript arguments
#postscript("bulkRNAseq_PCA_allsamples_labeledS5.eps")                           # Start graphics device driver
pca_norm_plot + ggtitle("PCA Normalized Counts -- Subset") +
  geom_label_repel(data=filter(all_df,
                               labels %in% c('s4_het_scn_s12', 's4_het_ret', 's4_het_scn', 's5_het_dlgn')),
                   nudge_x = 0.15,
                   nudge_y = 0.1)

pca_norm_plot + ggtitle("PCA Normalized Counts")# +

pca_norm_plot
# geom_label_repel(data=filter(all_df,
#                               labels %in% labels), #c('s4_het_scn', 's5_het_ret', 's5_het_scn', 's5_het_dlgn')),
#                   nudge_x = 0.15,
#                   nudge_y = 0.15)
#dev.off()
```

There really doesn't seem to be a discernible difference globally
between the hets and kos within each location. We can do a DE with
these samples between het/kos to quantify this.


# DE of KOs versus Low expressor Hets

```{r}
mm38_filt_sub <- subset_se(mm38_filt, subset = "sampleid != 'S3_Het_dLGN'&sampleid != 'S3_Het_Ret'&sampleid != 'S3_Het_SCN'&sampleid != 'S4_Het_dLGN' & sampleid != 'S4_Het_Ret' & sampleid != 'S4_Het_SCN'")

mm38_dlgn_subset <- subset_se(mm38_filt_sub, subset = "location_time == 'dLGN_het' | location_time == 'dLGN_ko'")
mm38_ret_subset <- subset_se(mm38_filt_sub, subset = "location_time == 'Retina_het' | location_time == 'Retina_ko'")
mm38_scn_subset <- subset_se(mm38_filt_sub, subset = "location_time == 'SCN_het' | location_time == 'SCN_ko'")

mm38_dlgn_subset <- normalize(mm38_dlgn_subset, filter = TRUE)
mm38_scn_subset <- normalize(mm38_scn_subset, filter = TRUE)
mm38_ret_subset <- normalize(mm38_ret_subset, filter = TRUE)

mm38_dlgn_subset <- set_batches(mm38_dlgn_subset, fact = "condition")
mm38_scn_subset <- set_batches(mm38_scn_subset, fact = "condition")
mm38_ret_subset <- set_batches(mm38_ret_subset, fact = "condition")

mm38_dlgn_subset <- set_conditions(mm38_dlgn_subset, fact = "location_time")
mm38_scn_subset <- set_conditions(mm38_scn_subset, fact = "location_time")
mm38_ret_subset <- set_conditions(mm38_ret_subset, fact = "location_time")
```

```{r}
test <- set_conditions(mm38_filt, fact = "location_time") |>
normalize(transform = "log2", convert = "cpm", batch = "sva")
plot_pca(test)
```

I do not think one should really do the DE with the data separate like
this unless there is some compelling reason.  I am going to repeat
without this constraint and see how similar/different the results are.
In addition, the current revision of all_pairwise() is much more
comfortable with more complex models, even with sva; so that might
make this more fun.

```{r}
mm_de_dlgn <- all_pairwise(mm38_dlgn_subset, model_fstring = "~ 0 + condition",
                           model_svs = "svaseq", filter = TRUE)
mm_de_dlgn
mm_table_dlgn <- combine_de_tables(mm_de_dlgn,
                                   excel = "39bulk_tables_tx/mm_de_dlgn_table.xlsx")
mm_sig_dlgn <- extract_significant_genes(mm_table_dlgn,
                                         excel = "39bulk_tables_tx/mm_de_dlgn_sig.xlsx")

mm_de_ret <- all_pairwise(mm38_ret_subset, model_fstring = "~ 0 + condition",
                          model_svs = "svaseq", filter = TRUE)
mm_de_ret
mm_table_ret <- combine_de_tables(mm_de_ret,
                                  excel = "39bulk_tables_tx/mm_de_ret_table.xlsx")
mm_sig_ret <- extract_significant_genes(mm_table_ret,
                                         excel = "39bulk_tables_tx/mm_de_ret_sig.xlsx")

mm_de_scn <- all_pairwise(mm38_scn_subset, model_fstring = "~ 0 + condition",
                          model_svs = "svaseq", filter = TRUE)
mm_de_scn
mm_table_scn <- combine_de_tables(mm_de_scn, excel = "39bulk_tables_tx/mm_de_scn_table.xlsx")
mm_sig_scn <- extract_significant_genes(mm_table_scn, according_to = "deseq",
                                        excel = "39bulk_tables_tx/mm_de_scn_sig.xlsx")
```

## SCN

```{r}
keepers <- names(mm_de_scn$deseq$all_tables)

res_tbls_scn <- c()
for (tbl in keepers) {
  deseq_tbl <- get_de_table(mm_de_scn, table = tbl)
  res_tbls_scn[[tbl]] <- merge(deseq_tbl, hisat_annot, by = "row.names")
  fctr <- c(paste(gsub("\\_.*","",tbl), "\nEnriched"),
            paste(sub('.*\\_', '', tbl), "\nEnriched"))
  res_tbls_scn[[tbl]] <- set_sig_limma(res_tbls_scn[[tbl]], factors = fctr)
  res_tbls_scn[[paste0(tbl, "volc_plotly")]] <- volc_plot(res_tbls_scn[[tbl]],
                                                          title = paste("Volcano Plot:", tbl),
                                                          type = "plotly",
                                                          tbl = "limma",
                                                          gene_name_col = "uniprot_gn_symbol")
  res_tbls_scn[[paste0(tbl, "volc_ggplot")]] <- volc_plot(res_tbls_scn[[tbl]],
                                                          title = paste("Volcano Plot:", tbl),
                                                          tbl = "limma")
}

res_tbls_scn$SCNko_vs_SCNhetvolc_ggplot
res_tbls_scn$SCNko_vs_SCNhetvolc_plotly
res_tbls_scn$SCNko_vs_SCNhet |>
  filter(adj.P.Val <= 0.05) |>
  filter(abs(logFC) >= 1) |>
  select("ensembl_gene_id", "uniprot_gn_symbol", "Significance", "description", "baseMean", "logFC", "adj.P.Val") |>
  arrange(logFC) |>
  DT::datatable(rownames = FALSE)
```

## dLGN

```{r}
keepers <- names(mm_de_dlgn$deseq$all_tables)

res_tbls_dlgn <- c()
for (tbl in keepers) {
  deseq_tbl <- get_de_table(mm_de_dlgn, table = tbl)
  res_tbls_dlgn[[tbl]] <- merge(deseq_tbl, hisat_annot, by = "row.names")
  res_tbls_dlgn[[tbl]] <- set_sig_limma(res_tbls_dlgn[[tbl]], factors = c(paste(gsub("\\_.*","",tbl), "\nEnriched"),
                                                                          paste(sub('.*\\_', '', tbl), "\nEnriched")))
  res_tbls_dlgn[[paste0(tbl, "volc_plotly")]] <- volc_plot(res_tbls_dlgn[[tbl]],
                                                           title = paste("Volcano Plot:", tbl),
                                                           type = "plotly",
                                                           tbl = "limma",
                                                           gene_name_col = "uniprot_gn_symbol")
  res_tbls_dlgn[[paste0(tbl, "volc_ggplot")]] <- volc_plot(res_tbls_dlgn[[tbl]],
                                                           title = paste("Volcano Plot:", tbl),
                                                           tbl = "limma")
}

res_tbls_dlgn$dLGNko_vs_dLGNhetvolc_ggplot
res_tbls_dlgn$dLGNko_vs_dLGNhetvolc_plotly
res_tbls_dlgn$dLGNko_vs_dLGNhet |>
  filter(adj.P.Val <= 0.05) |>
  filter(abs(logFC) >= 1) |>
  select("ensembl_gene_id", "uniprot_gn_symbol", "Significance", "description", "baseMean", "logFC", "adj.P.Val") |>
  arrange(logFC) |>
  DT::datatable(rownames = FALSE)
```


## Retina

```{r}
keepers <- names(mm_de_ret$deseq$all_tables)

res_tbls_ret <- c()
for (tbl in keepers) {
  deseq_tbl <- get_de_table(mm_de_ret, table = tbl)
  res_tbls_ret[[tbl]] <- merge(deseq_tbl, hisat_annot, by = "row.names")
  res_tbls_ret[[tbl]] <- set_sig_limma(res_tbls_ret[[tbl]], factors = c(paste(gsub("\\_.*","",tbl), "\nEnriched"),
                                                                        paste(sub('.*\\_', '', tbl), "\nEnriched")))
  res_tbls_ret[[paste0(tbl, "volc_plotly")]] <- volc_plot(res_tbls_ret[[tbl]],
                                                          title = paste("Volcano Plot:", tbl),
                                                          type = "plotly",
                                                          tbl = "limma",
                                                          gene_name_col = "uniprot_gn_symbol")
  res_tbls_ret[[paste0(tbl, "volc_ggplot")]] <- volc_plot(res_tbls_ret[[tbl]],
                                                          title = paste("Volcano Plot:", tbl),
                                                          tbl = "limma")
}

res_tbls_ret$Retinako_vs_Retinahetvolc_ggplot
res_tbls_ret$Retinako_vs_Retinahetvolc_plotly
res_tbls_ret$Retinako_vs_Retinahet |>
  filter(adj.P.Val <= 0.05) |>
  filter(abs(logFC) >= 1) |>
  select("ensembl_gene_id", "uniprot_gn_symbol", "Significance", "description", "baseMean", "logFC", "adj.P.Val") |>
  arrange(logFC) |>
  DT::datatable(rownames = FALSE)
```

# DE of all samples with SVA batch correction for high v low expressors

```{r}
mm38_dlgn_subset <- subset_se(
  mm38_filt, subset = "location_time == 'dLGN_het'| location_time == 'dLGN_ko'")
mm38_ret_subset <- subset_se(
  mm38_filt, subset = "location_time == 'Retina_het'| location_time == 'Retina_ko'")
mm38_scn_subset <- subset_se(
  mm38_filt, subset = "location_time == 'SCN_het'| location_time == 'SCN_ko'")

mm38_dlgn_subset <- normalize(mm38_dlgn_subset, filter = TRUE)
mm38_scn_subset <- normalize(mm38_scn_subset, filter = TRUE)
mm38_ret_subset <- normalize(mm38_ret_subset, filter = TRUE)

mm38_dlgn_subset <- set_batches(mm38_dlgn_subset, fact = "condition")
mm38_scn_subset <- set_batches(mm38_scn_subset, fact = "condition")
mm38_ret_subset <- set_batches(mm38_ret_subset, fact = "condition")

mm38_dlgn_subset <- set_conditions(mm38_dlgn_subset, fact = "location_time")
mm38_scn_subset <- set_conditions(mm38_scn_subset, fact = "location_time")
mm38_ret_subset <- set_conditions(mm38_ret_subset, fact = "location_time")
```

```{r}
mm_de_dlgn_sva <- all_pairwise(mm38_dlgn_subset, model_fstring = "~ 0 + condition",
                               model_svs = "svaseq", filter = TRUE)
mm_de_ret_sva <- all_pairwise(mm38_ret_subset, model_fstring = "~ 0 + condition",
                               model_svs = "svaseq", filter = TRUE)
mm_de_scn_sva <- all_pairwise(mm38_scn_subset, model_fstring = "~ 0 + condition",
                               model_svs = "svaseq", filter = TRUE)
mm_de_scn_sva

mm_scn_sva_table <- combine_de_tables(
  mm_de_scn_sva, excel = "excel/mm_subset_scn_sva_table.xlsx")
mm_scn_sva_table
mm_scn_sva_sig <- extract_significant_genes(
  mm_scn_sva_table, excel = "excel/mm_subset_scn_sva_sig.xlsx")
mm_scn_sva_sig
```

## dLGN

### PCA pre and post SVAseq

```{r}
mm38_dlgn_subset_norm_batch <- normalize(mm38_dlgn_subset, convert = "cpm",
                                         transform = "log2", batch = "svaseq")
mm38_dlgn_subset_norm_nobatch <- normalize(mm38_dlgn_subset, convert = "cpm", norm = "quant",
                                           transform = "log2")

pca_mm38_dlgn_subset_norm_nobatch <- plot_pca(mm38_dlgn_subset_norm_nobatch)
pca_norm_plot_nobatch <-  ggplot(pca_mm38_dlgn_subset_norm_nobatch$table, aes(x = PC1, y = PC2, label = labels)) +
  geom_point(aes(color = condition),
             size = 5) +
  scale_color_manual(values = color_values,
                     aesthetics = "color") +
  theme_classic() +
  labs(color="Cell Compartment/Genotype")   +
  ggtitle("PCA with No Batch Correction")  +
  labs(x = paste0("PC1: ",  pca_mm38_dlgn_subset_norm_nobatch$prop_var[1], "% variance"),
       y =  paste0("PC2: ",  pca_mm38_dlgn_subset_norm_nobatch$prop_var[2], "% variance"))  +
  geom_label_repel(data=filter(pca_mm38_dlgn_subset_norm_nobatch$table,
                               labels %in% c('S3_Ht_dLGN', 'S4_Ht_dLGN')),
                   nudge_x = 0.1,
                   nudge_y = 0.1)


pca_mm38_dlgn_subset_norm_batch <- plot_pca(mm38_dlgn_subset_norm_batch)
pca_norm_plot_batch <-  ggplot(pca_mm38_dlgn_subset_norm_batch$table, aes(x = PC1, y = PC2, label = labels)) +
  geom_point(aes(color = condition),
             size = 5) +
  scale_color_manual(values = color_values,
                     aesthetics = "color") +
  theme_classic() +
  labs(color="Cell Compartment/Genotype")   +
  ggtitle("PCA with SVA Batch Correction")  +
  labs(x = paste0("PC1: ",  pca_mm38_dlgn_subset_norm_batch$prop_var[1], "% variance"),
       y =  paste0("PC2: ",  pca_mm38_dlgn_subset_norm_batch$prop_var[2], "% variance"))  +
  geom_label_repel(data=filter(pca_mm38_dlgn_subset_norm_batch$table,
                               labels %in% c('S3_Ht_dLGN', 'S4_Ht_dLGN')),
                   nudge_x = 0.1,
                   nudge_y = 0.1)


pca_norm_plot_nobatch
pca_norm_plot_batch
```


### Volcano Plots ad DE table results

```{r}
keepers <- names(mm_de_dlgn_sva$deseq$all_tables)

res_tbls_dlgn_sva <- c()
for (tbl in keepers) {
  deseq_tbl <- get_de_table(mm_de_dlgn_sva, table = tbl)
  res_tbls_dlgn_sva[[tbl]] <- merge(deseq_tbl, hisat_annot, by = "row.names")
  res_tbls_dlgn_sva[[tbl]] <- set_sig_limma(res_tbls_dlgn_sva[[tbl]], factors = c(paste(gsub("\\_.*","",tbl), "\nEnriched"),
                                                                                  paste(sub('.*\\_', '', tbl), "\nEnriched")))
  res_tbls_dlgn_sva[[paste0(tbl, "volc_plotly")]] <- volc_plot(res_tbls_dlgn_sva[[tbl]],
                                                               title = paste("Volcano Plot:", tbl),
                                                               type = "plotly",
                                                               tbl = "limma",
                                                               gene_name_col = "uniprot_gn_symbol")
  res_tbls_dlgn_sva[[paste0(tbl, "volc_ggplot")]] <- volc_plot(res_tbls_dlgn_sva[[tbl]],
                                                               title = paste("Volcano Plot:", tbl),
                                                               tbl = "limma")
}

res_tbls_dlgn_sva$dLGNko_vs_dLGNhetvolc_ggplot
res_tbls_dlgn_sva$dLGNko_vs_dLGNhetvolc_plotly

res_tbls_dlgn_sva$dLGNko_vs_dLGNhet |>
  filter(adj.P.Val <= 0.05) |>
  filter(abs(logFC) >= 1) |>
  select("ensembl_gene_id", "uniprot_gn_symbol", "Significance", "description", "baseMean", "logFC", "adj.P.Val") |>
  arrange(logFC) |>
  DT::datatable(rownames = FALSE)
```

Samd11 and Dmrt3 are still both upregulated in dLGN KOs (as they were
without including the high-expressors).

Pttg1, Sema3g, Ifih1, Dppa5a, and Prg4 are all upregulated in dLGN
Hets including and excluding the high expressors. We do lose a few
genes when including high expressors that were enriched in dLGN hets
before.

## SCN

### PCA pre and post SVAseq

```{r}
mm38_scn_subset_norm_batch <- normalize(mm38_scn_subset, convert = "cpm",
                                        transform = "log2", batch = "svaseq")
mm38_scn_subset_norm_nobatch <- normalize(mm38_scn_subset, convert = "cpm", norm = "quant",
                                          transform = "log2")

pca_mm38_scn_subset_norm_nobatch <- plot_pca(mm38_scn_subset_norm_nobatch)
pca_norm_plot_nobatch_scn <-  ggplot(pca_mm38_scn_subset_norm_nobatch$table,
                                     aes(x = PC1, y = PC2, label = labels)) +
  geom_point(aes(color = condition),
             size = 5) +
  scale_color_manual(values = color_values[5:6],
                     aesthetics = "color") +
  theme_classic() +
  labs(color="Cell Compartment/Genotype")   +
  ggtitle("PCA with No Batch Correction")  +
  labs(x = paste0("PC1: ",  pca_mm38_scn_subset_norm_nobatch$prop_var[1], "% variance"),
       y =  paste0("PC2: ",  pca_mm38_scn_subset_norm_nobatch$prop_var[2], "% variance"))  +
  geom_label_repel(data=filter(pca_mm38_scn_subset_norm_nobatch$table,
                               labels %in% c('S3_Het_SCN', 'S4_Het_SCN')),
                   nudge_x = 0.2,
                   nudge_y = 0.2)


pca_mm38_scn_subset_norm_batch <- plot_pca(mm38_scn_subset_norm_batch)
pca_norm_plot_batch_scn <-  ggplot(pca_mm38_scn_subset_norm_batch$table, aes(x = PC1, y = PC2, label = labels)) +
  geom_point(aes(color = condition),
             size = 5) +
  scale_color_manual(values = color_values[5:6],
                     aesthetics = "color") +
  theme_classic() +
  labs(color="Cell Compartment/Genotype")   +
  ggtitle("PCA with SVA Batch Correction")  +
  labs(x = paste0("PC1: ",  pca_mm38_scn_subset_norm_batch$prop_var[1], "% variance"),
       y =  paste0("PC2: ",  pca_mm38_scn_subset_norm_batch$prop_var[2], "% variance"))  +
  geom_label_repel(data=filter(pca_mm38_scn_subset_norm_batch$table,
                               labels %in% c('S3_Het_SCN', 'S4_Het_SCN')),
                   nudge_x = 0.2,
                   nudge_y = 0.2)


pca_norm_plot_nobatch_scn
pca_norm_plot_batch_scn
```

### Volcano Plots ad DE table results

```{r}
keepers <- names(mm_de_scn_sva$deseq$all_tables)

res_tbls_scn_sva <- c()
for (tbl in keepers) {
  deseq_tbl <- get_de_table(mm_de_scn_sva, table = tbl)
  res_tbls_scn_sva[[tbl]] <- merge(deseq_tbl, hisat_annot, by = "row.names")
  res_tbls_scn_sva[[tbl]] <- set_sig_limma(res_tbls_scn_sva[[tbl]], factors = c(paste(gsub("\\_.*","",tbl), "\nEnriched"),
                                                                                paste(sub('.*\\_', '', tbl), "\nEnriched")))
  res_tbls_scn_sva[[paste0(tbl, "volc_plotly")]] <- volc_plot(res_tbls_scn_sva[[tbl]],
                                                              title = paste("Volcano Plot:", tbl),
                                                              type = "plotly",
                                                              tbl = "limma",
                                                              gene_name_col = "uniprot_gn_symbol")
  res_tbls_scn_sva[[paste0(tbl, "volc_ggplot")]] <- volc_plot(res_tbls_scn_sva[[tbl]],
                                                              title = paste("Volcano Plot:", tbl),
                                                              tbl = "limma")
}

res_tbls_scn_sva$SCNko_vs_SCNhetvolc_ggplot
res_tbls_scn_sva$SCNko_vs_SCNhetvolc_plotly

res_tbls_scn_sva$SCNko_vs_SCNhet |>
  filter(adj.P.Val <= 0.05) |>
  filter(abs(logFC) >= 1) |>
  select("ensembl_gene_id", "uniprot_gn_symbol", "Significance", "description", "baseMean", "logFC", "adj.P.Val") |>
  arrange(logFC) |>
  DT::datatable(rownames = FALSE)
```

## Retina

### PCA pre and post SVAseq

```{r}
mm38_ret_subset_norm_batch <- normalize(mm38_ret_subset, convert = "cpm",
                                        transform = "log2", batch = "svaseq")
mm38_ret_subset_norm_nobatch <- normalize(mm38_ret_subset, convert = "cpm", norm = "quant",
                                          transform = "log2")

pca_mm38_ret_subset_norm_nobatch <- plot_pca(mm38_ret_subset_norm_nobatch)
pca_norm_plot_nobatch_ret <-  ggplot(pca_mm38_ret_subset_norm_nobatch$table, aes(x = PC1, y = PC2, label = labels)) +
  geom_point(aes(color = condition),
             size = 5) +
  scale_color_manual(values = color_values[3:4],
                     aesthetics = "color") +
  theme_classic() +
  labs(color="Cell Compartment/Genotype")   +
  ggtitle("PCA with No Batch Correction")  +
  labs(x = paste0("PC1: ",  pca_mm38_ret_subset_norm_nobatch$prop_var[1], "% variance"),
       y =  paste0("PC2: ",  pca_mm38_ret_subset_norm_nobatch$prop_var[2], "% variance"))  +
  geom_label_repel(data=filter(pca_mm38_ret_subset_norm_nobatch$table,
                               labels %in% c('S3_Het_Ret', 'S4_Het_Ret')),
                   nudge_x = 0.2,
                   nudge_y = 0.2)


pca_mm38_ret_subset_norm_batch <- plot_pca(mm38_ret_subset_norm_batch)
pca_norm_plot_batch_ret <-  ggplot(pca_mm38_ret_subset_norm_batch$table, aes(x = PC1, y = PC2, label = labels)) +
  geom_point(aes(color = condition),
             size = 5) +
  scale_color_manual(values = color_values[3:4],
                     aesthetics = "color") +
  theme_classic() +
  labs(color="Cell Compartment/Genotype")   +
  ggtitle("PCA with SVA Batch Correction")  +
  labs(x = paste0("PC1: ",  pca_mm38_ret_subset_norm_batch$prop_var[1], "% variance"),
       y =  paste0("PC2: ",  pca_mm38_ret_subset_norm_batch$prop_var[2], "% variance"))  +
  geom_label_repel(data=filter(pca_mm38_ret_subset_norm_batch$table,
                               labels %in% c('S3_Het_Ret', 'S4_Het_Ret')),
                   nudge_x = 0.2,
                   nudge_y = 0.2)


pca_norm_plot_nobatch_ret
pca_norm_plot_batch_ret
```

### Volcano Plots ad DE table results

```{r}
keepers <- names(mm_de_ret_sva$deseq$all_tables)

res_tbls_ret_sva <- c()
for (tbl in keepers) {
  deseq_tbl <- get_de_table(mm_de_ret_sva, table = tbl)
  res_tbls_ret_sva[[tbl]] <- merge(deseq_tbl, hisat_annot, by = "row.names")
  res_tbls_ret_sva[[tbl]] <- set_sig_limma(res_tbls_ret_sva[[tbl]], factors = c(paste(gsub("\\_.*","",tbl), "\nEnriched"),
                                                                                paste(sub('.*\\_', '', tbl), "\nEnriched")))
  res_tbls_ret_sva[[paste0(tbl, "volc_plotly")]] <- volc_plot(res_tbls_ret_sva[[tbl]],
                                                              title = paste("Volcano Plot:", tbl),
                                                              type = "plotly",
                                                              tbl = "limma",
                                                              gene_name_col = "uniprot_gn_symbol")
  res_tbls_ret_sva[[paste0(tbl, "volc_ggplot")]] <- volc_plot(res_tbls_ret_sva[[tbl]],
                                                              title = paste("Volcano Plot:", tbl),
                                                              tbl = "limma")
}

res_tbls_ret_sva$Retinako_vs_Retinahetvolc_ggplot
res_tbls_ret_sva$Retinako_vs_Retinahetvolc_plotly

res_tbls_ret_sva$Retinako_vs_Retinahet |>
  filter(adj.P.Val <= 0.05) |>
  filter(abs(logFC) >= 1) |>
  select("ensembl_gene_id", "uniprot_gn_symbol", "Significance", "description", "baseMean", "logFC", "adj.P.Val") |>
  arrange(logFC) |>
  DT::datatable(rownames = FALSE)
```

# Re-Implementation

Let us start with the full set of desired tables.  In the above, the
data was separated into ~ 9 separate chunks and I think it will
benefit from being kept together -- also I seem to recall expressing
this to Theresa and that she changed the notebook to this kind of
analysis.  If I am correct, then it is possible she did so in a
separate notebook which I have not found and that this will provide
the fully reconciled set of results.

I have reasonable hope for this because the various plots etc early on
look identical between this containerized notebook, and those
notebooks I found in her working tree.

While I wait for these tables to come back, let us count the number of
samples in the plot provided by Colenso so that I may doublecheck that
nothing got filtered out or anything important like that.

8 pink samples, 17 green samples, 17 blue samples.

4 light pink, 4 dark pink
9 medium green, 4 light green, 4 dark green.
9 medium blue, 4 light blue, 4 dark blue.

With the above in mind, let us make a small change in how the data is
handled and see if I cannot get numbers more closely aligned to the
tables from Colenso.

```{r}
## Here is a reminder of Theresa's filter:
## filter(sample_id %in% c("S3_Het_Ret_S1", "S4_Het_Ret_S4", "S5_Het_Ret_S16", "S1_Het_Ret_S1", "S2_Het_Ret_S4", "S3_Het_Ret_S7", "S4_Het_Ret_S10", "S5_Het_Ret_S13")) |>

mm_final <- subset_se(mm38_hisat, subset = "included=='yes'") |>
  set_conditions(fact = "location_genotype")

mm_norm <- normalize(mm_final, transform = "log2", convert = "cpm", filter = TRUE, norm = "quant")
plot_pca(mm_norm)

mm_nb <- normalize(mm_final, transform = "log2", convert = "cpm", filter = TRUE, batch = "svaseq")
plot_pca(mm_nb)
```

There remain a couple of weirdo samples, but things look quite a bit
easier to interpret to my eyes.

```{r}
final_keepers <- list(
  ## Three comparisons among the dLGN samples.
  "dlgn_kw" = c("dLGN_ko", "dLGN_wt"),
  "dlgn_kh" = c("dLGN_ko", "dLGN_het"),
  "dlgn_hw" = c("dLGN_het", "dLGN_wt"),
  ## Three comparisons among the SCN samples.
  "scn_kw" = c("SCN_ko", "SCN_wt"),
  "scn_kh" = c("SCN_ko", "SCN_het"),
  "scn_hw" = c("SCN_het", "SCN_wt"),
  ## and the retina samples
  "retina_kw" = c("Retina_ko", "Retina_wt"),
  "retina_kh" = c("Retina_ko", "Retina_het"),
  "retina_hw" = c("Retina_het", "Retina_wt")
)

genoloc_de_sva <- all_pairwise(mm_final, model_fstring = "~ 0 + condition", filter = TRUE,
                               model_svs = "svaseq", keepers = final_keepers)
genoloc_table_sva <- combine_de_tables(genoloc_de_sva, excel = "excel/genotype_location_tables.xlsx")
genoloc_sig_sva <- extract_significant_genes(genoloc_table_sva, according_to = "deseq",
                                             excel = "excel/genotype_location_sig.xlsx")


location_keepers <- list(
  ## ko across locations
  "ko_dlgn_scn" = c("dLGN_ko", "SCN_ko"),
  "ko_dlgn_ret" = c("dLGN_ko", "Retina_ko"),
  "ko_scn_ret" = c("SCN_ko", "Retina_ko"),
  ## het across locations
  "het_dlgn_scn" = c("dLGN_het", "SCN_het"),
  "het_dlgn_ret" = c("dLGN_het", "Retina_het"),
  "het_scn_ret" = c("SCN_het", "Retina_het"),
  ## wt across locations
  "wt_dlgn_scn" = c("dLGN_wt", "SCN_wt"),
  "wt_dlgn_ret" = c("dLGN_wt", "Retina_wt"),
  "wt_scn_ret" = c("SCN_wt", "Retina_wt"))
location_comparisons <- all_pairwise(mm_final, model_fstring = "~ 0 + condition", filter = TRUE,
                               model_svs = "svaseq", keepers = location_keepers)
location_table <- combine_de_tables(location_comparisons, excel = "excel/locations.xlsx")
location_sig <- extract_significant_genes(location_table, according_to = "deseq",
                                             excel = "excel/location_sig.xlsx")
```

Save some of the DE results for comparison agains the TRAP data

```{r, eval=FALSE}
de_list <- c("mm_de_dlgn_sva", "mm_de_ret_sva", "mm_de_scn_sva",
             "mm_de_dlgn", "mm_de_ret", "mm_de_scn",
             "genoloc_de_sva", "genoloc_table_sva")
saved <- save(list = de_list, file = "bulk_results.rda")
```

