---
title: "ipRGC TRAP vs. bulk -- Compare the trap and bulk results."
output:
  html_document:
    code_download: true
    code_folding: hide
    fig_caption: true
    fig_height: 7
    fig_width: 7
    highlight: tango
    keep_md: false
    mode: selfcontained
    number_sections: true
    self_contained: true
    theme: readable
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
      toc_depth: 5
      rmdformats::readthedown:
        code_download: true
        code_folding: show
        df_print: paged
        fig_caption: true
        fig_height: 7
        fig_width: 7
        highlight: tango
        width: 300
        keep_md: false
        mode: selfcontained
        toc_float: true
        BiocStyle::html_document:
          code_download: true
          code_folding: show
          fig_caption: true
          fig_height: 7
          fig_width: 7
          highlight: tango
          keep_md: false
          mode: selfcontained
          toc_float: true
---
<style type="text/css">
body, td {
  font-size: 16px;
}
code.r{
  font-size: 16px;
}
pre {
  font-size: 16px
}
</style>

# Introduction

Collect the results from the gene/tx TRAP and bulk analyses here and
perform a few AUCC/correlation/etc comparisons to see how
similar/different they are.

*** Reminder to myself ***

When I did this on my main computer, I had the same environment loaded
for the bulk and trap data.  In the container this is not possible; so
I will need to judiciously use the rda argument to combine_de_tables()
and collect up the results before running the following blocks.  Don't
forget to add that.

# Compare the bulk and TRAP data

There are a few variables from the bulk analyses which we may use to
compare to the TRAP data:

mm_de_dlgn_sva, mm_de_ret_sva, mm_de_scn_sva, mm_de_dlgn, mm_de_ret,
mm_de_scn.

Those are location-specific contrasts.

genoloc_de_sva/genoloc_table_sva provide all contrasts that I thought
might be fun.

The following is all pre-wt filtering.  Next I will repeat everything post-filtering.

```{r, eval=FALSE}
## The bulk tables
genoloc_table_sva[["table_names"]]

## The TRAP tables
location_tables_full[["table_names"]]
genotype_tables_full[["table_names"]]
time_tables_full[["table_names"]]

## Retina het vs Retina wt
bulk_table <- genoloc_table_sva[["data"]][["Retina_ko_vs_Retina_het"]]
trap_table <- genotype_tables_full[["data"]][["kh_p08_retina"]]
retina_ko_het_aucc <- calculate_aucc(bulk_table, trap_table,
                                     px = "deseq_adjp", py = "deseq_adjp",
                                     lx = "deseq_logfc", ly = "deseq_logfc")
retina_ko_het_aucc
retina_ko_het_ros <- rank_order_scatter(bulk_table, trap_table)
retina_ko_het_ros[["lfc_scatter"]]

bulk_table <- genoloc_table_sva[["data"]][["Retina_ko_vs_Retina_het"]]
trap_table <- genotype_tables_full[["data"]][["kh_p15_retina"]]
retina_ko_het_aucc <- calculate_aucc(bulk_table, trap_table,
                                     px = "deseq_adjp", py = "deseq_adjp",
                                     lx = "deseq_logfc", ly = "deseq_logfc")
retina_ko_het_aucc
retina_ko_het_ros <- rank_order_scatter(bulk_table, trap_table)
retina_ko_het_ros[["lfc_scatter"]]

bulk_table <- genoloc_table_sva[["data"]][["SCN_ko_vs_SCN_het"]]
trap_table <- genotype_tables_full[["data"]][["kh_p08_scn"]]
scn_ko_het_aucc <- calculate_aucc(bulk_table, trap_table,
                                     px = "deseq_adjp", py = "deseq_adjp",
                                     lx = "deseq_logfc", ly = "deseq_logfc")
scn_ko_het_aucc
scn_ko_het_ros <- rank_order_scatter(bulk_table, trap_table)
scn_ko_het_ros[["lfc_scatter"]]


bulk_table <- genoloc_table_sva[["data"]][["SCN_ko_vs_SCN_het"]]
trap_table <- genotype_tables_full[["data"]][["kh_p15_scn"]]
scn_ko_het_aucc <- calculate_aucc(bulk_table, trap_table,
                                     px = "deseq_adjp", py = "deseq_adjp",
                                     lx = "deseq_logfc", ly = "deseq_logfc")
scn_ko_het_aucc
scn_ko_het_ros <- rank_order_scatter(bulk_table, trap_table)
scn_ko_het_ros[["lfc_scatter"]]

bulk_table <- genoloc_table_sva[["data"]][["dLGN_ko_vs_dLGN_het"]]
trap_table <- genotype_tables_full[["data"]][["kh_p08_dlgn"]]
dlgn_ko_het_aucc <- calculate_aucc(bulk_table, trap_table,
                                     px = "deseq_adjp", py = "deseq_adjp",
                                     lx = "deseq_logfc", ly = "deseq_logfc")
dlgn_ko_het_aucc
dlgn_ko_het_ros <- rank_order_scatter(bulk_table, trap_table)
dlgn_ko_het_ros[["lfc_scatter"]]

bulk_table <- genoloc_table_sva[["data"]][["dLGN_ko_vs_dLGN_het"]]
trap_table <- genotype_tables_full[["data"]][["kh_p15_dlgn"]]
dlgn_ko_het_aucc <- calculate_aucc(bulk_table, trap_table,
                                     px = "deseq_adjp", py = "deseq_adjp",
                                     lx = "deseq_logfc", ly = "deseq_logfc")
dlgn_ko_het_aucc
dlgn_ko_het_ros <- rank_order_scatter(bulk_table, trap_table)
dlgn_ko_het_ros[["lfc_scatter"]]
```

```{r, eval=FALSE}
## The bulk tables
genoloc_table_sva[["table_names"]]

## The TRAP tables
location_tables[["table_names"]]
genotype_tables[["table_names"]]
time_tables[["table_names"]]

## Retina het vs Retina wt
bulk_table <- genoloc_table_sva[["data"]][["Retina_ko_vs_Retina_het"]]
trap_table <- genotype_tables_full[["data"]][["kh_p08_retina"]]
retina_ko_het_aucc <- calculate_aucc(bulk_table, trap_table,
                                     px = "deseq_adjp", py = "deseq_adjp",
                                     lx = "deseq_logfc", ly = "deseq_logfc")
retina_ko_het_aucc
retina_ko_het_ros <- rank_order_scatter(bulk_table, trap_table)
retina_ko_het_ros[["lfc_scatter"]]

bulk_table <- genoloc_table_sva[["data"]][["Retina_ko_vs_Retina_het"]]
trap_table <- genotype_tables_full[["data"]][["kh_p15_retina"]]
retina_ko_het_aucc <- calculate_aucc(bulk_table, trap_table,
                                     px = "deseq_adjp", py = "deseq_adjp",
                                     lx = "deseq_logfc", ly = "deseq_logfc")
retina_ko_het_aucc
retina_ko_het_ros <- rank_order_scatter(bulk_table, trap_table)
retina_ko_het_ros[["lfc_scatter"]]

bulk_table <- genoloc_table_sva[["data"]][["SCN_ko_vs_SCN_het"]]
trap_table <- genotype_tables_full[["data"]][["kh_p08_scn"]]
scn_ko_het_aucc <- calculate_aucc(bulk_table, trap_table,
                                     px = "deseq_adjp", py = "deseq_adjp",
                                     lx = "deseq_logfc", ly = "deseq_logfc")
scn_ko_het_aucc
scn_ko_het_ros <- rank_order_scatter(bulk_table, trap_table)
scn_ko_het_ros[["lfc_scatter"]]


bulk_table <- genoloc_table_sva[["data"]][["SCN_ko_vs_SCN_het"]]
trap_table <- genotype_tables_full[["data"]][["kh_p15_scn"]]
scn_ko_het_aucc <- calculate_aucc(bulk_table, trap_table,
                                     px = "deseq_adjp", py = "deseq_adjp",
                                     lx = "deseq_logfc", ly = "deseq_logfc")
scn_ko_het_aucc
scn_ko_het_ros <- rank_order_scatter(bulk_table, trap_table)
scn_ko_het_ros[["lfc_scatter"]]

bulk_table <- genoloc_table_sva[["data"]][["dLGN_ko_vs_dLGN_het"]]
trap_table <- genotype_tables_full[["data"]][["kh_p08_dlgn"]]
dlgn_ko_het_aucc <- calculate_aucc(bulk_table, trap_table,
                                     px = "deseq_adjp", py = "deseq_adjp",
                                     lx = "deseq_logfc", ly = "deseq_logfc")
dlgn_ko_het_aucc
dlgn_ko_het_ros <- rank_order_scatter(bulk_table, trap_table)
dlgn_ko_het_ros[["lfc_scatter"]]

bulk_table <- genoloc_table_sva[["data"]][["dLGN_ko_vs_dLGN_het"]]
trap_table <- genotype_tables_full[["data"]][["kh_p15_dlgn"]]
dlgn_ko_het_aucc <- calculate_aucc(bulk_table, trap_table,
                                     px = "deseq_adjp", py = "deseq_adjp",
```
