---
title: "An initial exploration of the IPRGC re-processed samples."
author: "atb abelew@gmail.com"
date: "`r Sys.Date()`"
output:
  html_document:
    code_download: true
    code_folding: show
    fig_caption: true
    fig_height: 7
    fig_width: 7
    highlight: zenburn
    keep_md: false
    mode: selfcontained
    number_sections: true
    self_contained: true
    theme: readable
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
  rmdformats::readthedown:
    code_download: true
    code_folding: show
    df_print: paged
    fig_caption: true
    fig_height: 7
    fig_width: 7
    highlight: zenburn
    width: 300
    keep_md: false
    mode: selfcontained
    toc_float: true
  BiocStyle::html_document:
    code_download: true
    code_folding: show
    fig_caption: true
    fig_height: 7
    fig_width: 7
    highlight: zenburn
    keep_md: false
    mode: selfcontained
    toc_float: true
---

<style type="text/css">
body, td {
  font-size: 16px;
}
code.r{
  font-size: 16px;
}
pre {
  font-size: 16px
}
body .main-container {
  max-width: 1600px;
}
</style>

```{r options, include=FALSE}
library(hpgltools)
library(ggplot2)
library(ggstatsplot)
tt <- try(devtools::load_all("~/hpgltools"))
knitr::opts_knit$set(
  progress = TRUE, verbose = TRUE, width = 90, echo = TRUE)
knitr::opts_chunk$set(
  error = TRUE, fig.width = 8, fig.height = 8, fig.retina = 2,
  out.width = "100%", dev = "png",
  dev.args = list(png = list(type = "cairo-png")))
old_options <- options(digits = 4, stringsAsFactors = FALSE, knitr.duplicate.label = "allow")
ggplot2::theme_set(ggplot2::theme_bw(base_size = 12))
ver <- "202406"
previous_file <- ""
ver <- format(Sys.Date(), "%Y%m%d")

##tmp <- sm(loadme(filename=paste0(gsub(pattern="\\.Rmd", replace="", x=previous_file), "-v", ver, ".rda.xz")))
rmd_file <- "initial_analyses.Rmd"
```

# Meeting with Theresa

Previous papers did not do an explicit subtraction, instead just
compared to WT and kept the genes which are > in delta/het vs. wt.
There are multiple ways to deal with this and that query has not yet
been defined.  Later, Theresa came to the conclusion that the
subtraction method is not appropriate.

# Introduction

In this document I hope to explore the freshly processed samples and
perform some comparisons to see that we have the expected similarities
and differences from the prior analysis performed by Theresa.

There is one way in which I expect any/all of these analyses to be
explicitly different: this should include the changes produced by
April's renaming of some samples.

My intention is to produce a sample sheet which includes one column
with non-umi-deduplicated results and one with deduplicated results.
With the exception of the previous point, I hope that the first will
be identical (or at least very close to identical) to Theresa's result
while the second I expect will be subtly different -- but I am hoping
subtly enough that it will not significantly change the interpretation
but be a little more precise.

Lets see!  I need therefore to make a change to my metadata gathering
function to include the umi deduplicated result.  I am thinking
therefore to create a separate specification for umi-barcoded samples
because looking through the logs for umi stuff when they are not used
will be too much of a pain...

```{r}
umi_spec <- make_rnaseq_spec(umi = TRUE)
iprgc_2022_meta <- gather_preprocessing_metadata("sample_sheets/20240606_only_umd_sequenced.xlsx",
                                                 spec = umi_spec, species = "mm39_112", verbose = FALSE,
                                                 basedir = "preprocessing/umd_sequenced")
```

From this point on, I am hoping/intending to pull liberally from
Theresa's notebook with a diversion to compare the three datasets:

* Pre-April renaming: E.g. Theresa's current dataset
* Post renaming: Unless I am mistaken, this should be very similar to
  the above.
* Post deduplication: Given what I saw from the extracted logs in the
  sample sheet, I expect this to be similar but not identical to the
  previous two.

Lets find out!   But first, annotations!

# Annotation data

I am pulling this from Theresa's anxontrapR_pipeline.Rmd, primarily
because it looks similar to the other documents, but was modified more
recently.  I will change it slightly, primarily because I grabbed a
new mmusculus assembly and therefore I will pull the mmusculus
annotations from a specific biomart archive that should match it.

```{r}
mm_annot <- load_biomart_annotations(species = "mmusculus", year = "2023", month = "02")
mm_annot <- mm_annot[["annotation"]]
mm_annot[["txid"]] <- paste0(mm_annot[["ensembl_transcript_id"]], ".", mm_annot[["version"]])
rownames(mm_annot) <- make.names(mm_annot[["ensembl_gene_id"]], unique=TRUE)
tx_gene_map <- mm_annot[, c("txid", "ensembl_gene_id")]
```

# Hisat2 expressionsets

The primary difference between my block and Theresa's are:

1.  I am pulling the metadata directly from the
    gather_preprocessing_metadata() above.
2.  I am using the column 'symlink' which is just a copy of the
    existing 'file' column with a small change so I can load it from
    my directory without having to copy everything.
3.  I am using the ensembl genome release 39, version 112 and so
    pulled a somewhat newer copy of the annotation data.
4.  The original is named 'v1', followed by 'v2' and 'v3' for the
    other two treatments I performed.

## Color choices

Given that we are excluding a bunch of the older samples, the set of
colors I expect to find is different; so I will make explicit here the
various colors used to denote location/genotype/time/etc.

April turned me onto this website 'paletton.com' for this kind of
stuff and I will try and pick out palettes which basically match what
I am getting with the original colors.

```{r}
color_choices <- list(
  "geno_loc" = list(
    "het_dlgn" = "#9ECAE1",
    "het_retina" = "#F46D43",
    "het_scn" = "#2CA25F",
    "ko_dlgn" = "#3182BD",
    "ko_retina" = "#FDAE61",
    "ko_scn" = "#006D2C",
    "wt_dlgn" = "#DEEBF7",
    "wt_retina" = "#D73027",
    "wt_scn" = "#66C2A4"),
  ## These colors are coming from ipRGC_summaryplots.html
  ## I am using kcolorchooser to grab them rather than get confused by the text
  "location" = list(
    "retina" = "#d73027",
    "dlgn" = "#3182bd",
    "scn" = "#006b29"),
  "genotype" = list(
    "wt" = "#D4D4D4",
    "het" = "#787878",
    "ko" = "#313131"),
  "time" = list(
    "p08" = "#5E104B",
    "p15" = "#4E9231"))
```

There is one noteworthy sample: iprgc_103, it was effectively replaced
when April renamed the samples and so exists in the v1 data, but not
v2/v3; they instead have the newly named samples which I called
iprgc_123 to iprgc_130.  As a result, I copied the annotations for
iprgc_123 to my column so that there is no discrepency in terms of
genotype/location/time.

```{r}
mm38_hisat_v1 <- create_expt(iprgc_2022_meta[["new_meta"]],
                             gene_info = mm_annot,
                             file_column = "symlink") %>%
  set_expt_conditions(fact = "genolocatb") %>%
  set_expt_batches(fact = "timeatb") %>%
  set_expt_colors(color_choices[["geno_loc"]])
mm38_hisat_v1

mm38_hisat_v2 <- create_expt(iprgc_2022_meta[["new_meta"]], gene_info = mm_annot,
                             file_column = "hisat_count_table") %>%
  set_expt_conditions(fact = "genolocatb") %>%
  set_expt_batches(fact = "timeatb") %>%
  set_expt_colors(color_choices[["geno_loc"]])
mm38_hisat_v2

mm38_hisat_v3 <- create_expt(iprgc_2022_meta[["new_meta"]], gene_info = mm_annot,
                             file_column = "umi_dedup_output_count") %>%
  set_expt_conditions(fact = "genolocatb") %>%
  set_expt_batches(fact = "timeatb") %>%
  set_expt_colors(color_choices[["geno_loc"]])
mm38_hisat_v3
```

# Non-zero Counts per Sample

Let's look at the number of non-zero genes for all samples versus the
coverage.

```{r}
plot_legend(mm38_hisat_v1)

v1_nonzero <- plot_nonzero(mm38_hisat_v1)
v1_nonzero

v2_nonzero  <- plot_nonzero(mm38_hisat_v2)
v2_nonzero

v3_nonzero  <- plot_nonzero(mm38_hisat_v3)
v3_nonzero
```

Oh wow, I did not expect such a profound effect on the cpm values on
the more saturated libraries.  I guess in retrospect I should have?

Also note to self, we are not messing with p60.

```{r}
mm38_hisat_v1 <- subset_expt(mm38_hisat_v1, subset = "timeatb!='p60'")
mm38_hisat_v2 <- subset_expt(mm38_hisat_v2, subset = "timeatb!='p60'")
mm38_hisat_v3 <- subset_expt(mm38_hisat_v3, subset = "timeatb!='p60'")
```

# Quick PCA, then return to Theresa's document

```{r}
v1_norm <- normalize_expt(mm38_hisat_v1, transform = "log2", convert = "cpm",
                          norm = "quant", filter = TRUE)
v2_norm <- normalize_expt(mm38_hisat_v2, transform = "log2", convert = "cpm",
                          norm = "quant", filter = TRUE)
v3_norm <- normalize_expt(mm38_hisat_v3, transform = "log2", convert = "cpm",
                          norm = "quant", filter = TRUE)

plot_pca(v1_norm)
plot_pca(v2_norm)
plot_pca(v3_norm)
```

To my eyes it looks like we just have 1 weirdo p15 sample?
Deduplication had a minor but significant effect on the PCA.

With that in mind, let us look at Theresa's WORKING document and see
what we can recapitulate.

Theresa's document: The TRAP protocol has some variability which is
introduced at different steps including homogenization, antibody
labeling, pulldown efficiency/specificity, sample handling during
cleanup steps, and library prep/sequencing. We know from Rashmi's QC
that there is variability at the level of pulldown efficiency (amount
of RNA isolated). She is doing a good job of keeping track of this for
all her samples and we have validated her P8 results (attached
supplementary figure 3D). We consistently see clear differences
between control and cre samples for the retina, which makes sense
because the cell bodies are in the retina. The target tissue
differences are smaller, which also makes sense for axon-TRAP. We
think that some of her P15 samples are not good based on low amounts
of isolated RNA from cre(+) retina samples. We plan to drop these
samples and not perform additional isolations at this time
point. Based on this (and the general lack of large developmental
effects), we were planning to focus on presenting the P8 data only in
the paper. Interested to hear your thoughts in this...

My notes: Theresa's first operations in this notebook were to:

1.  Set location as condition, genotype as batch.
2.  Perform PCA before/after sva.

```{r}
v3_loc_geno <- set_expt_conditions(mm38_hisat_v3, fact = "locationatb",
                                   colors = color_choices[["location"]]) %>%
  set_expt_batches(fact = "genotypeatb")
```

## The associated PCA

At different times, it appears to me that Theresa has preferred
slightly different normalization methods, primarily a mix of TMM and
quantile.

Thus I will use different suffix letters to denote various
normalizations employed, and if they turn out the same I will pick one arbitrarily.

```{r}
loc_geno_nq <- normalize_expt(v3_loc_geno, transform = "log2", convert = "cpm",
                              filter = TRUE, norm = "quant")
plot_pca(loc_geno_nq)
## ok, I have two weirdo samples which look very much like they are actually dlgn.
## These are sample IDs iprgc_66 and iprgc_130

loc_geno_nt <- normalize_expt(v3_loc_geno, transform = "log2", convert = "cpm",
                             filter = TRUE, norm = "tmm")
plot_pca(loc_geno_nt)
```

A random thought about these PCA plots, it might be worth while to add
a panel below the legend with the sample numbers per condition/batch.

Of course, the same information is provided in a more fun fashion via
my silly sankey function:

```{r}
sample_sankey <- plot_meta_sankey(v3_loc_geno, color_choices = color_choices,
                                  factors = c("genotypeatb", "locationatb", "timeatb"))
sample_sankey
```

# A Short conversation with Rashmi

Rashmi came by and we discussed the samples a little.  She suggested
that is likely that we will need to exclude the 202205 samples, these
may be identified by a few ways, most easily I think via the
'projectah' column, they are the 021_1 samples.

My sense was that she concurred with my interpretation of the umi
deduplication, so I will continue using the deduplicated results
exclusively, at least for now.

# Melanopsin Sanity Check

One of Theresa's first checks was wisely for melanopsin.  Let us
repeat a version of this:

```{r}
opn4_exprs <- data.frame(combined = pData(loc_geno_nt)[["genolocatb"]],
                         location = pData(loc_geno_nt)[["locationatb"]],
                         genotype = pData(loc_geno_nt)[["genotypeatb"]],
                         opn = exprs(loc_geno_nt)["ENSMUSG00000021799", ])

groupedstats::grouped_summary(opn4_exprs, location, opn)
ggbetweenstats(data = opn4_exprs, x = location, y = opn)
ggbetweenstats(data = opn4_exprs, x = genotype, y = opn)
ggbetweenstats(data = opn4_exprs, x = combined, y = opn)
```

ok, so I plotted the question a bit differently, but got the same
answer.

Here is the text of Theresa's notebook following this analysis:

"Ugh oh, looks like there is at least one retina KO sample that has
some melanopsin expression in it. Turns out ipRGC_07 is a bad egg
which is supposed to be a KO but has melanopsin expression. It’s
friends which were pooled from the same mice are iprgc_06 and
iprgc_08, so we need to exclude all these samples."

I am also seeing some knockout expression with some caveats: I do not
have the affected samples in my dataset (iprgc_07) and the levels I am
seeing are quite low -- I will look in IGV to double check, but I
strongly suspect that these are some piddly reads near the UTRs.

Onward!

# Library sizes post-deduplication

Theresa's next operation was to perform libsize/nonzero plots.  I
already did the pre/post deduplication nonzero, here is the analagous
libsize.

v2 is pre-deduplication and v3 is post.

```{r}
plot_libsize(mm38_hisat_v2)
plot_libsize(mm38_hisat_v3)
```

I am a bit concerned about some of these library sizes
post-deduplication.

Let us look at the relationship between reads and duplication, which I
assume will be relatively linear.

```{r}
test <- pData(mm38_hisat_v3)[, c("hisatgenomesingleall", "umideduppctreads")]
test_plot <- plot_linear_scatter(test)
test_plot[["scatter"]]
```

Theresa also produced a density/sample plot, that might prove quite
useful for these due to their significantly larger variance across
samples (due to deduplication).

```{r}
plot_density(mm38_hisat_v3)
```

There is some difference across sample densities, but it is not too
crazytown.

# PCA plots

## PCA of all genes by location

Theresa's first pca was of log2 cpm values.  I might add quantile/tmm
to this?

```{r}
v3_location <- set_expt_conditions(mm38_hisat_v3, fact = "locationatb") %>%
  set_expt_batches(fact = "genotypeatb")

v3_location_norm <- normalize_expt(v3_location, filter = TRUE, norm = "quant",
                                   transform = "log2", convert = "cpm")
plot_pca(v3_location_norm)
```

Once again we see that samples iprgc_66 and iprgc_130 are likely
actually DLGN and not SCN.  I am therefore going to add a column to
the sample sheet noting this, and remove them from the expressionset.

I will thus replot the data after removing those two.  If we want to
see what it looks like with the re-attributed locations, we can do so.

Theresa has a nice change to the PCA plotter in which she sets the
alpha channel as an additional visual queue for a metadata factor...

```{r}
mm38_hisat_v3 <- subset_expt(mm38_hisat_v3, subset="sampleid!='iprgc_130'") %>%
  subset_expt(subset="sampleid!='iprgc_66'")
v3_location <- set_expt_conditions(mm38_hisat_v3, fact = "locationatb") %>%
  set_expt_batches(fact = "genotypeatb")
v3_location_norm <- normalize_expt(v3_location, filter = TRUE, norm = "quant",
                                   transform = "log2", convert = "cpm")
plot_pca(v3_location_norm)

removed_sankey <- plot_meta_sankey(v3_location, color_choices = color_choices,
                                   factors = c("genotypeatb", "locationatb", "timeatb"))
removed_sankey
```

Here is Theresa's text, recall once again that I do not have some of
these older samples (iprgc_62):

PC1 vs PC2 identifies retina vs axon is still the main component of
variation. We do see though that in the PC2 direction, we see with the
new samples added, we don’t see separation based on axonal targets
(dLGN vs SCN). In the PC1 vs PC3 plot, we see that it’s PC3 where we
start to see variation correlated with axonal compartment. Let’s look
at PC1 vs PC2 colored by batch (when they were processed/sequenced) to
see if that is what is contributing so much variation in PC2.

Side note: ipRGC 62 seems like an odd ball. This seems to me like it
should have been a dLGN P08 sample. Is there any possibility this got
mislabeled early on? I went back and double checked to see if all my
processing is correct and it indeed was labeled an SCN P15 from the
time I got the samples, and it is indeed.

# DE

I now switched to Theresa's document 'WORKING_axonTRAP...' and will
start pulling sections from it.  I am reasonably certain I have
reasonably similar sample distributions, so I presume I can invoke
similar/identical calls for DESeq and friends.

## p8 retinas

In the block immediately before the DE analyses, Theresa created a
subset expressionset of only p08 retinas.  Thus this initial DE I
assume will be used to subtract for the SCN/DLGN analyses that follow.
(I guess I could read ahead and find out, but no! I want to be a
blank slate)

```{r}
mm38_p8_retina <- subset_expt(mm38_hisat_v3, subset = "timeatb=='p08' & locationatb=='retina'")

mm_normal_p8_ret_de <- all_pairwise(mm38_p8_retina, model_batch = "svaseq", filter = TRUE)
mm_normal_p8_ret_de
```

The following invocation performed by Theresa filters the wt/het
comparison for only those genes which increased by at least 0.25 logFC
with a significant adjusted p-value.  I assume that this is to use the
wt samples as a translational control for the ket/ko comparisons; I am
therefore thinking that for my purposes, I will therefore separate the
contrasts from all_pairwise do this in a stepwise fashion...

The block of code immediately following Theresa's all_pairwise()
invocation is a little confusing for me and warrants some explanation
by me to me in the hopes that I do not misunderstand what is happening
and the goals therein.

Here is Theresa's next line:

```{r, eval=FALSE}
hetkeeper_genes <- mm_de_normal_p8_ret$deseq$all_tables$wtretina_vs_hetretina %>%
  filter(logFC <= -.25 & adj.P.Val <= 0.05)
```

I think I can safely assume that the goal here is to pull out the IDs
which increased in het with respect to wild type; even if by a small
margin, as long as it is statistically significant vis a vis the
adjusted p-value.

I am going to perform what I think is the same thing in a slightly
different fashion so that I can share a copy of the results with
whomever is interested.  I will also repeat Theresa's invocation and
prove to myself that I understood and got the same answer.

```{r}
wt_het_keeper <- list("het_vs_wt" = c("hetretina", "wtretina"))
het_wt_table <- combine_de_tables(mm_normal_p8_ret_de, keepers = wt_het_keeper,
                                  excel = "excel/het_retina_control.xlsx")

wanted_sig <- extract_significant_genes(het_wt_table,
                                        lfc = 0.25,
                                        according_to = "deseq")

wanted_het_increased <- wanted_sig[["deseq"]][["ups"]][["het_vs_wt"]]
increased_het_genes <- rownames(wanted_het_increased)
```

## Prove I understood

```{r}
hetkeeper_genes <- mm_normal_p8_ret_de$deseq$all_tables$wtretina_vs_hetretina %>%
  filter(logFC <= -0.25 & adj.P.Val <= 0.05)
testthat::expect_true(nrow(hetkeeper_genes) == length(increased_het_genes))
```

Yay! I can read!  Now let us repeat for the KO vs wt

```{r}
wt_ko_keeper <- list("ko_vs_wt" = c("koretina", "wtretina"))
ko_wt_table <- combine_de_tables(mm_normal_p8_ret_de, keepers = wt_ko_keeper,
                                 excel = "excel/ko_retina_control.xlsx")
wanted_sig <- extract_significant_genes(ko_wt_table,
                                        lfc = 0.25,
                                        according_to = "deseq")
wanted_ko_increased <- wanted_sig[["deseq"]][["ups"]][["ko_vs_wt"]]
increased_ko_genes <- rownames(wanted_ko_increased)
```

The next thing performed in Theresa's document is a unique(concatenation of
these two gene groups), thus sucking up every gene which was
significantly higher in either the knockout _or_ heterzyous samples
with respect to wild-type.

This was followed by a couple of merge operations of a little bit of
the annotation data; I am not sure I understand the goal yet...

Here is her code:

```{r, eval=FALSE}
keepergenes <- unique(c(rownames(hetkeeper_genes), rownames(kokeeper_genes)))

annots_to_merge <- mm_annot %>%
    select(ensembl_gene_id, external_gene_name) %>%
    filter(ensembl_gene_id %in% rownames(mm_de_normal_p8_ret$deseq$all_tables$koretina_vs_hetretina)) %>%
    distinct()

mm_de_normal_p8_ret$deseq$all_tables$koretina_vs_hetretina <- merge(mm_de_normal_p8_ret$deseq$all_tables$koretina_vs_hetretina, annots_to_merge,
                                                                by.x = 0,
                                                                by.y = "ensembl_gene_id",
                                                                all.x = TRUE)

df <- mm_de_normal_p8_ret$deseq$all_tables$koretina_vs_hetretina %>%
    dplyr::mutate(logFC = -logFC) %>%
    set_sig_limma(factors = c("Het Enriched",
                              "KO Enriched"))
```

```{r}
both_increased_genes <- unique(c(increased_het_genes, increased_ko_genes))
```
