---
title: "IPRGC translatome analyses: `r Sys.getenv('VERSION')`: README."
author: "atb abelew@gmail.com"
bibliography: /home/trey/scratch/zotero_library/atb.bib
date: "`r Sys.Date()`"
output:
  html_document:
    code_download: true
    code_folding: show
    fig_caption: true
    fig_height: 7
    fig_width: 7
    highlight: zenburn
    keep_md: false
    mode: selfcontained
    number_sections: true
    self_contained: true
    theme: readable
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
  rmdformats::readthedown:
    code_download: true
    code_folding: show
    df_print: paged
    fig_caption: true
    fig_height: 7
    fig_width: 7
    highlight: zenburn
    width: 300
    keep_md: false
    mode: selfcontained
    toc_float: true
  BiocStyle::html_document:
    code_download: true
    code_folding: show
    fig_caption: true
    fig_height: 7
    fig_width: 7
    highlight: zenburn
    keep_md: false
    mode: selfcontained
    toc_float: true
---

<style type="text/css">
body, td {
  font-size: 16px;
}
code.r {
  font-size: 16px;
}
pre {
  font-size: 16px
}
body .main-container {
  max-width: 1600px;
}
</style>

```{r options, include=FALSE}
knitr::opts_knit$set(
  progress = TRUE, verbose = TRUE, width = 90, echo = TRUE)
knitr::opts_chunk$set(
  error = TRUE, fig.width = 8, fig.height = 8, fig.retina = 2,
  out.width = "100%", dev = "png",
  dev.args = list(png = list(type = "cairo-png")))
old_options <- options(digits = 4, stringsAsFactors = FALSE, knitr.duplicate.label = "allow")
ggplot2::theme_set(ggplot2::theme_bw(base_size = 12))
ver <- Sys.getenv("VERSION")
```

# TODO

# Changelog

# Introduction

This document is intended to provide some background into these
analyses, packaging them into a downloadable container, and any
notes/ideas that cross my mind.

The section titled 'Roadmap' is intended to provide a quick reference
to find any figures produced by these analyses.

# Packaging the counts

Currently, the container image for these analyses starts at the mapped
counts and progresses through the markdown documents to create the
html logs, figures, and xlsx documents.  The following block reads the
project metadata in order to create an archive file containing those
counts.  One might reasonably ask why it is not compressed by default, the
individual count tables are aggressively compressed, so that would be
redundant redundant.

```{r, eval=FALSE}
deduplicated_tarball <- tar_meta_column("sample_sheets/20240606_only_umd_sequenced_modified.xlsx",
                                        column = "umi_dedup_output_count", output = "deduplicated_counts.tar",
                                        compression = NULL)
original_tarball <- tar_meta_column("sample_sheets/20240606_only_umd_sequenced_modified.xlsx",
                                     column = "hisat_count_table", output = "non_deduplicated_counts.tar",
                                     compression = NULL)
symlink_tarball <-  tar_meta_column("sample_sheets/20240606_only_umd_sequenced_modified.xlsx",
                                    column = "symlink", output = "symlink_counts.tar",
                                    compression = NULL)
```



# Roadmap

# Bibliography
