---
title: "IPRGC translatome analyses: `r Sys.getenv('VERSION')`: README."
author: "atb abelew@gmail.com"
bibliography: /home/trey/scratch/zotero_library/atb.bib
date: "`r Sys.Date()`"
output:
  html_document:
    code_download: true
    code_folding: show
    fig_caption: true
    fig_height: 7
    fig_width: 7
    highlight: zenburn
    keep_md: false
    mode: selfcontained
    number_sections: true
    self_contained: true
    theme: readable
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
  rmdformats::readthedown:
    code_download: true
    code_folding: show
    df_print: paged
    fig_caption: true
    fig_height: 7
    fig_width: 7
    highlight: zenburn
    width: 300
    keep_md: false
    mode: selfcontained
    toc_float: true
  BiocStyle::html_document:
    code_download: true
    code_folding: show
    fig_caption: true
    fig_height: 7
    fig_width: 7
    highlight: zenburn
    keep_md: false
    mode: selfcontained
    toc_float: true
---

<style type="text/css">
body, td {
  font-size: 16px;
}
code.r {
  font-size: 16px;
}
pre {
  font-size: 16px
}
body .main-container {
  max-width: 1600px;
}
</style>

```{r options, include=FALSE}
knitr::opts_knit$set(
  progress = TRUE, verbose = TRUE, width = 90, echo = TRUE)
knitr::opts_chunk$set(
  error = TRUE, fig.width = 8, fig.height = 8, fig.retina = 2,
  out.width = "100%", dev = "png",
  dev.args = list(png = list(type = "cairo-png")))
old_options <- options(digits = 4, stringsAsFactors = FALSE, knitr.duplicate.label = "allow")
ggplot2::theme_set(ggplot2::theme_bw(base_size = 12))
ver <- Sys.getenv("VERSION")
```

# TODO

# Changelog

# Introduction

This document is intended to provide some background into these
analyses, packaging them into a downloadable container, and any
notes/ideas that cross my mind.

The section titled 'Roadmap' is intended to provide a quick reference
to find any figures produced by these analyses.

# Packaging the counts

Currently, the container image for these analyses starts at the mapped
counts and progresses through the markdown documents to create the
html logs, figures, and xlsx documents.  The following block reads the
project metadata in order to create an archive file containing those
counts.  One might reasonably ask why it is not compressed by default, the
individual count tables are aggressively compressed, so that would be
redundant redundant.

```{r, eval=FALSE}
deduplicated_tarball <- tar_meta_column("sample_sheets/20240606_only_umd_sequenced_modified.xlsx",
                                        column = "umi_dedup_output_count", output = "deduplicated_counts.tar",
                                        compression = NULL)
original_tarball <- tar_meta_column("sample_sheets/20240606_only_umd_sequenced_modified.xlsx",
                                     column = "hisat_count_table", output = "non_deduplicated_counts.tar",
                                     compression = NULL)
symlink_tarball <-  tar_meta_column("sample_sheets/20240606_only_umd_sequenced_modified.xlsx",
                                    column = "symlink", output = "symlink_counts.tar",
                                    compression = NULL)
```


# Output organization

This iteration of the iprgc analyses adopted a peculiar organization
strategy.  Each set of images/xlsx/etc files produced goes into a
separate numeric prefixed directory.

I did this primarily in an attempt to make clear the distinction
between observations before removing genes higher in the wt
samples vs. after wt removal.

+ 01diagnostic_images: The set of plots taken before doing pretty much
  anything else.  This includes showing the differences between
  pre/post umi-deduplication, clustering before removing the
  problemat samples (either due to low coverage or seriously
  inappropriate sample clustering).  Note: all of plots of the 'v1'
  data are excluded; they are sufficiently redundant that I dropped
  them.
+ 02filtered_images: Mostly a repetition of the 01 directory, but now
  the final set of samples/genes which survived filtering.  If we were
  to use any of the diagnostic images as supplemental, they would be
  here. (none were)
+ 03theresa_comparison_excel: This document was intended as a simple
  transcription of Theresa's notebook before I started messing with
  it.  This directory is intended to trace back and show that the
  results I am getting are very similar to Theresa's -- which is true,
  but there are some differences introduced by the slightly different
  filtering performed and the usage of UMI deduplicated inputs.
+ In the notebook, after the 03 section, there is a large number of
  not-run blocks.  These are copy/pasted from Theresa's notebook and
  provide the starting point for my changes.  I am pretty sure they
  all still work and one is welcome to run them.
+ 04inclusion_comparisons: Every informative comparison against wt
  samples resides here.  We have in these files every gene which is
  highly observed in the wt samples, and as such, are expected to be
  the result of mRNAs non-specifically binding during the TRAP
  experiment.
+ 05inclusion_volcano_ma: Plots of the 04 results.  A quick and easy
  way to get a feeling for what was non specific.  We kind of
  hope to see a grab-bag of randomness here.
+ (hah I cannot count, there is no 06 directory), I think I had one
  with gprofiler of the wt stuff, but deleted it.
+ 07included_strict_genes_excel: A large xlsx file of all high-wt gene
  IDs.
+ 08full_contrasts_excel: The large xlsx output files comparing
  everything which is _not_ wt.  E.g. this has the heterozygous and
  knockout comparisons of genotype, location, and time.  No genes were
  excluded from these tables.
+ 09full_contrasts_intersections: The set of intersections of genes
  among the not-filtered comparisons.
+ 10genotype_contrasts_excel: This is effectively a copy of the 08
  files, but with only those genes higher in the het/ko samples than
  their corresponding wt samples.  E.g. this take the 08 results and
  keeps only the stuff in the 07 directory.
+ 11enrichment_excel: The various results from clusterProfiler and
  gProfiler2, these take the genes in the of significant (p <= 0.05,
  |lfc| >= 1.0 or whatever) set from the 10genotype_contrasts_excel
  directory and pass them to the over-enrichment and gene set analyses
  against a mix of GO, reactome, transcription-factors, miRNA, KEGG,
  mSigDB, etc...
+ 12clusterprofiler_plots: Plots from 11.
+ 13genotype_ma_volcano: MA and volcano plots of the ko/het
  comparisons in pdf format.  These include labels of specific genes
  of particular interest.
+ 14genotype_strict_contrasts_excel: A repetition of 10 above, but
  with a stricter definition of 'significant'.
+ 15: xlsx output from clusterProfiler/gProfiler2 of the results from
  14.
+ 16clusterProfiler_plots: Plots from 15.
+ 17full_location_contrasts: The tables comparing paired genotype
  sampletypes by location (retina, dlgn, scn).  These tables do not
  remove any genes observed higher in their corresponding wt.
+ 18location_contrasts: Post filtering copies of 17.
+ 19: The xlsx outputs of gProfiler2/clusterProfiler for 18.
+ 19clusterProfiler_plots: (I think this should be 20): Plots from 19.
+ 20location_ma_volcano: MA/volcano plots from 18.
+ 21location_strict_contrasts: Same as 14, but for the locations
  (stricter p-value/lfc).
+ 22: Ontology etc. results from 21.
+ 23clusterProfiler_plots: Plots from 22.
+ 24location_ma_volcano_strict: Same as 20, but stricter.
+ 25full_contrasts_time: Same as 17, but this time comparing across
  p08 and p15.
+ 26time_contrasts: Excluding the high-wt from 25.
+ 27: The ontology xlsx files from 26.
+ 28clusterProfiler_plots: The plots from 27.
+ 29time_ma_volcano: MA/volcano plots from 26.
+ 30time_strict_contrasts_excel: Repeating the strict fc/p call for
  the time comparisons.
+ 31: The ontology xlsx files from 30.
+ 32: Plots of 31.
+ 33translatome_xlsx: xlsx results of unfiltered a/b / c/d comparisons: the
  translatomes.  Note: there is a pre/post sva version of this because
  of the complexity of these contrasts.  This includes a peculiar
  subtraction I did using the deseq results.
+ 34translatome_deseqsub_xlsx: 33, but using the DESeq2 results to
  exclude genes.
+ 35translatome_union: Using unions/intersections to filter genes.
+ 36upset_genotype: Getting groups of genes by union/intersection from
  upset.
+ 37upset_location: Ibid but the locations.
+ 38msigdb: Using GSVA to compare the results against the mh/m1/etc
  gene sets from the mSigDB revision 2024.  I am not sure about the
  legality of this and may have removed it.

I tried out some random stuff after #38 but did not write the results.

# Roadmap

# Bibliography
