---
title: "ipRGC bulk RNAseq -- All Samples"
author: "Theresa Alexander"
date: "4/15/2023"
output:
  html_document:
    code_download: true
    code_folding: hide
    fig_caption: true
    fig_height: 7
    fig_width: 7
    highlight: tango
    keep_md: false
    mode: selfcontained
    number_sections: true
    self_contained: true
    theme: readable
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 5
  rmdformats::readthedown:
    code_download: true
    code_folding: show
    df_print: paged
    fig_caption: true
    fig_height: 7
    fig_width: 7
    highlight: tango
    width: 300
    keep_md: false
    mode: selfcontained
    toc_float: true
  BiocStyle::html_document:
    code_download: true
    code_folding: show
    fig_caption: true
    fig_height: 7
    fig_width: 7
    highlight: tango
    keep_md: false
    mode: selfcontained
    toc_float: true
---
<style type="text/css">
body, td {
  font-size: 16px;
}
code.r{
  font-size: 16px;
}
pre {
 font-size: 16px
}
</style>


```{r libraries, message=FALSE}
library("hpgltools")
library("ggplot2")
library("patchwork")
library("plotly")
library("RColorBrewer")
library("DT")
library("dplyr")
library("openxlsx")
library("ComplexHeatmap")
library("circlize")
library("gprofiler2")
source("helper_functions.R")
library("gplots")
library("ggrepel")
```


```{r}
#setwd("~/scratch_02/reesyxan/mmusculus_iprgc_2019")
```

# M. musculus

## Annotations

I am using mm38_100.

```{r annotations}
mm_annot <- load_biomart_annotations(species="mmusculus")
mm_annot <- mm_annot[["annotation"]]
mm_annot[["txid"]] <- paste0(mm_annot[["ensembl_transcript_id"]], ".", mm_annot[["version"]])
rownames(mm_annot) <- make.names(mm_annot[["ensembl_gene_id"]], unique=TRUE)

tx_gene_map <- mm_annot[, c("txid", "ensembl_gene_id")]
```


So, I now have a table of mouse annotations.

## Hisat2 expressionset

```{r hisat2}
hisat_annot <- mm_annot
rownames(hisat_annot) <- paste0(rownames(hisat_annot))
mm38_hisat <- create_expt("sample_sheets/samplesheet_bulkrnaseq_Opn4.xlsx", gene_info=hisat_annot)
```



# Color Set Up

```{r}
color_condition <- mm38_hisat$initial_metadata$condition
color_condition <- factor(color_condition, levels = c("Retina_het", "Retina_ko", "SCN_het", "SCN_ko", "dLGN_het", "dLGN_ko"))

color_values <- c( "#00BFC4",  "#9AD2D3", "#F8766D", "#FBADA7", "#7CAE00", "#D3EE8F")
col_fun = colorRamp2(c(0, 3, 6), c("blue", "pink", "red"))
```


# Mapped Counts Per Sample Plot

```{r}
library(forcats)
mm38_hisat <- set_expt_conditions(mm38_hisat, fact = "batch")
libsize_df <- as.data.frame(plot_libsize(mm38_hisat)$table)
rownames(libsize_df) <- mm38_hisat$design$sampleid
libsize_df$samplenum <- paste0(mm38_hisat$design[rownames(libsize_df), "samplenumb"], "_", 
       mm38_hisat$design[rownames(libsize_df), "batch"])
libsize_df$sample_id <-  rownames(mm38_hisat$design)
libsize_df <- libsize_df[order(libsize_df$sum),]


ggplot(libsize_df, aes(x = fct_rev(fct_infreq(samplenum, sum)), y = sum, fill = condition)) + 
    geom_bar(stat = "identity", color = "black") +
    theme_classic() +
    #theme(axis.text.x= element_text(fct_rev(fct_infreq(libsize_df$samplenum, libsize_df$sum))), #remove x axis labels
    #    axis.ticks.x=element_blank()) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    scale_fill_manual(values = c("#60449b", "#009f6f", "orange")) +
    xlab("Sample") +
    ylab("Sum of Mapped Gene Counts") +
    labs(fill = "")
```


# Normalize Expression

First we will filter out the genes which have low counts across all samples. Then I will normalize both without and with SVA correction.

```{r}
mm38_filt <- normalize_expt(mm38_hisat, filter=TRUE)
mm38_norm <- normalize_expt(mm38_filt, filter=TRUE, 
                             convert="cpm", 
                             norm = "quant", 
                             transform="log2")

mm38_norm <- set_expt_conditions(mm38_norm, fact = paste0(mm38_norm$design$location, "_", mm38_norm$design$genotype))
```
# Melanopsin Expression

Here is a plot of the Opn4 expression across all retina samples split by batch. The 4 quadrants in the plot are <br>
het:1 -- retina het samples in batch 1 <br>
het:2 -- retina het samples in batch 2 <br>
ko:1 -- retina ko samples in batch 1 <br>
ko:2 -- retina ko samples in batch 2 <br>



```{r}
colors <- brewer.pal(name = "Spectral", n = 2)
mm38_cpm <- normalize_expt(mm38_filt, filter=TRUE, 
                             convert="cpm")
df <- data.frame(exprs = exprs(mm38_cpm$expressionset)["ENSMUSG00000021799",], 
                 genotype = as.factor(mm38_cpm$design$genotype), 
                 location = as.factor(mm38_cpm$design$location), 
                 sampleid = mm38_cpm$design$sampleid,
                 samplenum = mm38_cpm$design$samplenumb,
                 batch = mm38_cpm$design$batch)

setEPS()                                             # Set postscript arguments
postscript("bulkRNAseq_Opn4Expression_Retina.eps")                           # Start graphics device driver
df %>%
  filter(location == "Retina" | location == "retina") %>%
  ggplot(aes(x=samplenum, y=exprs, fill=genotype, group = genotype)) +
 geom_bar(stat="identity", position = position_dodge(width = 0.8)) +
  xlab("Sample") +
  ylab("Opn4 Counts Per Million (CPMs)") + 
  #scale_fill_manual(values = c("#FFC107", "#004D40", "orange"), 
  #                  aesthetics = "fill") +
  ggtitle("Retina Melanopsin Expression") +
     facet_wrap(~genotype:batch, scales="free") +
  ylim(c(0,25)) +
  theme_classic()
dev.off() 


```


The WT samples in batch 3 have about double the Opn4 expression as the het samples in batch 3 (and batch 2). 

# PCA

```{r pca, message = FALSE}
pca_norm <- plot_pca(mm38_norm, plot_labels = FALSE)
labels <- paste0(mm38_norm$design$samplenumb, "_", mm38_norm$design$batch)#rownames(pca_norm$result$v)
#labels <- paste0(mm38_norm$design$sampleid, "_", mm38_norm$design$samplenumb)
all_df <-  cbind(as.data.frame(pca_norm$result$v), labels)

shape_genotype <- mm38_norm$design$genotype
color_condition <- mm38_norm$condition


# filter(sample_id %in% c("S3_Het_Ret_S1", "S4_Het_Ret_S4", "S5_Het_Ret_S16", "S1_Het_Ret_S1", "S2_Het_Ret_S4", "S3_Het_Ret_S7", "S4_Het_Ret_S10", "S5_Het_Ret_S13")) %>%


pca_norm_plot <-  ggplot(all_df, aes(x = PC1, y = PC2, label = labels)) + 
  geom_point(aes(color = color_condition, shape = shape_genotype), 
                      size = 5) +
  #scale_color_manual(values = color_values, 
  #                  aesthetics = "color") + 
  theme_classic() +
  labs(color="Cell Compartment/Genotype")   +    
  ggtitle("PCA with No Batch Correction")  +
  labs(x = paste0("PC1: ",  pca_norm$prop_var[1], "% variance"), 
       y =  paste0("PC2: ",  pca_norm$prop_var[2], "% variance")) 

#setEPS()                                             # Set postscript arguments
#postscript("bulkRNAseq_PCA_allsamples_labeledS5.eps")                           # Start graphics device driver
pca_norm_plot + ggtitle("PCA Normalized Counts") #+ 
  #geom_label_repel(data=filter(all_df, 
   #                            labels %in% labels), #c('s5_het_dlgn', 's5_het_ret', 's5_het_scn')),
               #    nudge_x = 0.15,
              #     nudge_y = 0.15)
#dev.off() 
```



```{r}
#batch 1 plot
all_df[rownames(all_df) %in% tolower(rownames(mm38_norm$design[mm38_norm$design$batch == "1",])),] %>%
  ggplot(aes(x = PC1, y = PC2)) + 
  geom_point(aes(color = color_condition[which(rownames(all_df) %in% tolower(rownames(mm38_norm$design[mm38_norm$design$batch == "1",])))], 
                 shape = shape_genotype[which(rownames(all_df) %in% tolower(rownames(mm38_norm$design[mm38_norm$design$batch == "1",])))]), 
                      size = 5) +
  scale_color_manual(values = color_values, 
                    aesthetics = "color") + 
  theme_classic() +
  labs(color="Cell Compartment/Genotype", shape = "Genotype")   +    
  ggtitle("PCA - Batch 1 RNAseq")  +
  labs(x = paste0("PC1: ",  pca_norm$prop_var[1], "% variance"), 
       y =  paste0("PC2: ",  pca_norm$prop_var[2], "% variance")) 



#batch 1&2 plot
pca_batch1and2 <- all_df %>%
  ggplot(aes(x = PC1, y = PC2)) + 
  geom_point(aes(color = color_condition, 
                 shape = shape_genotype), 
                      size = 5) +
  scale_color_manual(values = color_values, 
                    aesthetics = "color") + 
  theme_classic() +
  labs(color="Cell Compartment/Genotype", shape = "Genotype")   +    
  ggtitle("PCA - Batch 1&2 RNAseq")  +
  labs(x = paste0("PC1: ",  pca_norm$prop_var[1], "% variance"), 
       y =  paste0("PC2: ",  pca_norm$prop_var[2], "% variance")) 


#setEPS()
#postscript(file = "rnaseq_pca-batch1and2.eps")
#pca_batch1and2
#dev.off()
```
<br><br>

In this pca plot, we see the retina samples cluster on the right in the PC1 direction and the brain samples cluster on the left in the PC 1 direction. 
<br>
<br>
As we saw before, there is disparity between some of the hets. As it turns out, all of the new Het samples cluster with the KO samples as our one weirdo did before. So now it seems that it is quite easy to get some genetic drift that leads to these mice's expression to behave in weird ways. 
<br>
<br>
As we did before, let's look at the Opn4 expression across all the retina samples to see if the Het samples that behave like KOs have lower Opn4 expression. 
<br>
<br>



# A closer look at the similar KOs and Hets

Let's exclude the 2 Hets that are the most dissimilar from the bunch and look at a PCA plot of the rest of the samples.

```{r}
mm38_subset <- subset_expt(mm38_filt, subset = "sampleid != 'S3_Het_dLGN'&sampleid != 'S3_Het_Ret'&sampleid != 'S3_Het_SCN'&sampleid != 'S4_Het_dLGN'&sampleid != 'S4_Het_Ret'&sampleid != 'S4_Het_SCN'")

mm38_subset <- normalize_expt(mm38_subset, filter=TRUE, 
                             convert="cpm", 
                             norm = "quant", 
                             transform="log2")

mm38_subset <- set_expt_conditions(mm38_subset, fact = "locationtime")

pca_norm_subset <- plot_pca(mm38_subset, plot_labels = FALSE)

labels <- rownames(pca_norm_subset$result$v)

all_df <-  cbind(as.data.frame(pca_norm_subset$result$v), labels)
shape_genotype <- mm38_subset$design$genotype
color_condition <- pca_norm_subset$table$condition



pca_norm_plot <-  ggplot(pca_norm_subset$table, aes(x = PC1, y = PC2)) + 
  geom_point(aes(color = condition), 
                      size = 5) +
  scale_color_manual(values = color_values, 
                    aesthetics = "color") + 
  theme_classic() +
  labs(color="Cell Compartment/Genotype")   +    
  ggtitle("PCA with No Batch Correction")  +
  labs(x = paste0("PC1: ",  pca_norm_subset$prop_var[1], "% variance"), 
       y =  paste0("PC2: ",  pca_norm_subset$prop_var[2], "% variance")) 



pca_norm_plot <-  ggplot(all_df, aes(x = PC1, y = PC2, label = labels)) + 
  geom_point(aes(color = color_condition, shape = shape_genotype), 
                      size = 5) +
  #scale_color_manual(values = color_values, 
  #                  aesthetics = "color") + 
  theme_classic() +
  labs(color="Cell Compartment/Genotype")   +    
  ggtitle("PCA with No Batch Correction")  +
  labs(x = paste0("PC1: ",  pca_norm_subset$prop_var[1], "% variance"), 
       y =  paste0("PC2: ",  pca_norm_subset$prop_var[2], "% variance")) 

#setEPS()                                             # Set postscript arguments
#postscript("bulkRNAseq_PCA_allsamples_labeledS5.eps")                           # Start graphics device driver
pca_norm_plot + ggtitle("PCA Normalized Counts -- Subset") + 
  geom_label_repel(data=filter(all_df, 
                               labels %in% c('s4_het_scn_s12', 's4_het_ret', 's4_het_scn', 's5_het_dlgn')),
                   nudge_x = 0.15,
                   nudge_y = 0.1)

pca_norm_plot + ggtitle("PCA Normalized Counts")# + 
 # geom_label_repel(data=filter(all_df, 
#                               labels %in% labels), #c('s4_het_scn', 's5_het_ret', 's5_het_scn', 's5_het_dlgn')),
#                   nudge_x = 0.15,
#                   nudge_y = 0.15)
#dev.off() 
```

There really doesn't seem to be a discernible difference globally between the hets and kos within each location. We can do a DE with these samples between het/kos to quantify this.


# DE of KOs versus Low expressor Hets 

```{r}
mm38_filt_sub <- subset_expt(mm38_filt, subset = "sampleid != 'S3_Het_dLGN'&sampleid != 'S3_Het_Ret'&sampleid != 'S3_Het_SCN'&sampleid != 'S4_Het_dLGN' & sampleid != 'S4_Het_Ret' & sampleid != 'S4_Het_SCN'")

mm38_dlgn_subset <- subset_expt(mm38_filt_sub, subset = "locationtime == 'dLGN_het'| locationtime == 'dLGN_ko'")
mm38_ret_subset <- subset_expt(mm38_filt_sub, subset = "locationtime == 'Retina_het'| locationtime == 'Retina_ko'")
mm38_scn_subset <- subset_expt(mm38_filt_sub, subset = "locationtime == 'SCN_het'| locationtime == 'SCN_ko'")

mm38_dlgn_subset <- normalize_expt(mm38_dlgn_subset, filter = TRUE)
mm38_scn_subset <- normalize_expt(mm38_scn_subset, filter = TRUE)
mm38_ret_subset <- normalize_expt(mm38_ret_subset, filter = TRUE)

mm38_dlgn_subset <- set_expt_batches(mm38_dlgn_subset, fact = "condition")
mm38_scn_subset <- set_expt_batches(mm38_scn_subset, fact = "condition")
mm38_ret_subset <- set_expt_batches(mm38_ret_subset, fact = "condition")

mm38_dlgn_subset <- set_expt_conditions(mm38_dlgn_subset, fact = "locationtime")
mm38_scn_subset <- set_expt_conditions(mm38_scn_subset, fact = "locationtime")
mm38_ret_subset <- set_expt_conditions(mm38_ret_subset, fact = "locationtime")
```



```{r}
mm_de_dlgn <- all_pairwise(mm38_dlgn_subset, 
                           model_batch = "svaseq",
                           parallel=FALSE, 
                           do_ebseq=FALSE, 
                           do_limma=FALSE,
                           do_edger=FALSE)
    
mm_de_ret <- all_pairwise(mm38_ret_subset, 
                          model_batch = "svaseq",
                          parallel=FALSE, 
                          do_ebseq=FALSE, 
                          do_limma=FALSE,
                          do_edger=FALSE)

mm_de_scn <- all_pairwise(mm38_scn_subset, 
                          model_batch = "svaseq",
                          parallel=FALSE, 
                          do_ebseq=FALSE, 
                          do_limma=FALSE,
                          do_edger=FALSE)
```

## dLGN

```{r}
keepers <- names(mm_de_dlgn$deseq$all_tables)

res_tbls_dlgn <- c()
for (tbl in keepers){
  deseq_tbl <- mm_de_dlgn$deseq$all_tables[[tbl]]
  rownames(deseq_tbl) <- paste0("gene:", rownames(deseq_tbl))
  res_tbls_dlgn[[tbl]] <- merge(deseq_tbl, hisat_annot, by = "row.names")
  res_tbls_dlgn[[tbl]] <- set_sig_limma(res_tbls_dlgn[[tbl]], factors = c(paste(gsub("\\_.*","",tbl), "\nEnriched"), 
                                                                paste(sub('.*\\_', '', tbl), "\nEnriched")))
  res_tbls_dlgn[[paste0(tbl, "volc_plotly")]] <- volc_plot(res_tbls_dlgn[[tbl]], 
                                                      title = paste("Volcano Plot:", tbl), 
                                                      type = "plotly", 
                                                      tbl = "limma",
                                                      gene_name_col = "external_gene_name")
  res_tbls_dlgn[[paste0(tbl, "volc_ggplot")]] <- volc_plot(res_tbls_dlgn[[tbl]], 
                                                      title = paste("Volcano Plot:", tbl), 
                                                      tbl = "limma")
}

res_tbls_dlgn$dLGNko_vs_dLGNhetvolc_ggplot
res_tbls_dlgn$dLGNko_vs_dLGNhetvolc_plotly
res_tbls_dlgn$dLGNko_vs_dLGNhet %>%
    filter(adj.P.Val <= 0.05) %>%
    filter(abs(logFC) >= 1) %>%
    select("ensembl_gene_id", "external_gene_name", "Significance", "description", "baseMean", "logFC", "adj.P.Val") %>%
    arrange(logFC) %>%
    DT::datatable(rownames = FALSE)
```


## SCN

```{r}
keepers <- names(mm_de_scn$deseq$all_tables)

res_tbls_scn <- c()
for (tbl in keepers){
  deseq_tbl <- mm_de_scn$deseq$all_tables[[tbl]]
  rownames(deseq_tbl) <- paste0("gene:", rownames(deseq_tbl))
  res_tbls_scn[[tbl]] <- merge(deseq_tbl, hisat_annot, by = "row.names")
  res_tbls_scn[[tbl]] <- set_sig_limma(res_tbls_scn[[tbl]], factors = c(paste(gsub("\\_.*","",tbl), "\nEnriched"), 
                                                                paste(sub('.*\\_', '', tbl), "\nEnriched")))
  res_tbls_scn[[paste0(tbl, "volc_plotly")]] <- volc_plot(res_tbls_scn[[tbl]], 
                                                      title = paste("Volcano Plot:", tbl), 
                                                      type = "plotly", 
                                                      tbl = "limma",
                                                      gene_name_col = "external_gene_name")
  res_tbls_scn[[paste0(tbl, "volc_ggplot")]] <- volc_plot(res_tbls_scn[[tbl]], 
                                                      title = paste("Volcano Plot:", tbl), 
                                                      tbl = "limma")
}

res_tbls_scn$SCNko_vs_SCNhetvolc_ggplot
res_tbls_scn$SCNko_vs_SCNhetvolc_plotly
res_tbls_scn$SCNko_vs_SCNhet %>%
    filter(adj.P.Val <= 0.05) %>%
    filter(abs(logFC) >= 1) %>%
    select("ensembl_gene_id", "external_gene_name", "Significance", "description", "baseMean", "logFC", "adj.P.Val") %>%
    arrange(logFC) %>%
    DT::datatable(rownames = FALSE)
```




## Retina

```{r}
keepers <- names(mm_de_ret$deseq$all_tables)

res_tbls_ret <- c()
for (tbl in keepers){
  deseq_tbl <- mm_de_ret$deseq$all_tables[[tbl]]
  rownames(deseq_tbl) <- paste0("gene:", rownames(deseq_tbl))
  res_tbls_ret[[tbl]] <- merge(deseq_tbl, hisat_annot, by = "row.names")
  res_tbls_ret[[tbl]] <- set_sig_limma(res_tbls_ret[[tbl]], factors = c(paste(gsub("\\_.*","",tbl), "\nEnriched"), 
                                                                paste(sub('.*\\_', '', tbl), "\nEnriched")))
  res_tbls_ret[[paste0(tbl, "volc_plotly")]] <- volc_plot(res_tbls_ret[[tbl]], 
                                                      title = paste("Volcano Plot:", tbl), 
                                                      type = "plotly", 
                                                      tbl = "limma",
                                                      gene_name_col = "external_gene_name")
  res_tbls_ret[[paste0(tbl, "volc_ggplot")]] <- volc_plot(res_tbls_ret[[tbl]], 
                                                      title = paste("Volcano Plot:", tbl), 
                                                      tbl = "limma")
}

res_tbls_ret$Retinako_vs_Retinahetvolc_ggplot
res_tbls_ret$Retinako_vs_Retinahetvolc_plotly
res_tbls_ret$Retinako_vs_Retinahet %>%
    filter(adj.P.Val <= 0.05) %>%
    filter(abs(logFC) >= 1) %>%
    select("ensembl_gene_id", "external_gene_name", "Significance", "description", "baseMean", "logFC", "adj.P.Val") %>%
    arrange(logFC) %>%
    DT::datatable(rownames = FALSE)
```



# DE of all samples with SVA batch correction for high v low expressors


```{r}
mm38_dlgn_subset <- subset_expt(mm38_filt, subset = "locationtime == 'dLGN_het'| locationtime == 'dLGN_ko'")
mm38_ret_subset <- subset_expt(mm38_filt, subset = "locationtime == 'Retina_het'| locationtime == 'Retina_ko'")
mm38_scn_subset <- subset_expt(mm38_filt, subset = "locationtime == 'SCN_het'| locationtime == 'SCN_ko'")

mm38_dlgn_subset <- normalize_expt(mm38_dlgn_subset, filter = TRUE)
mm38_scn_subset <- normalize_expt(mm38_scn_subset, filter = TRUE)
mm38_ret_subset <- normalize_expt(mm38_ret_subset, filter = TRUE)

mm38_dlgn_subset <- set_expt_batches(mm38_dlgn_subset, fact = "condition")
mm38_scn_subset <- set_expt_batches(mm38_scn_subset, fact = "condition")
mm38_ret_subset <- set_expt_batches(mm38_ret_subset, fact = "condition")

mm38_dlgn_subset <- set_expt_conditions(mm38_dlgn_subset, fact = "locationtime")
mm38_scn_subset <- set_expt_conditions(mm38_scn_subset, fact = "locationtime")
mm38_ret_subset <- set_expt_conditions(mm38_ret_subset, fact = "locationtime")
```


```{r}
mm_de_dlgn_sva <- all_pairwise(mm38_dlgn_subset, 
                           model_batch = "svaseq",
                           parallel=FALSE, 
                           do_ebseq=FALSE, 
                           do_limma=FALSE,
                           do_edger=FALSE)
    
mm_de_ret_sva <- all_pairwise(mm38_ret_subset, 
                          model_batch = "svaseq",
                          parallel=FALSE, 
                          do_ebseq=FALSE, 
                          do_limma=FALSE,
                          do_edger=FALSE)

mm_de_scn_sva <- all_pairwise(mm38_scn_subset, 
                          model_batch = "svaseq",
                          parallel=FALSE, 
                          do_ebseq=FALSE, 
                          do_limma=FALSE,
                          do_edger=FALSE)
```

## dLGN

### PCA pre and post SVAseq

```{r}
mm38_dlgn_subset_norm_batch <- normalize_expt(mm38_dlgn_subset, convert = "cpm", transform = "log2", batch = "svaseq")
mm38_dlgn_subset_norm_nobatch <- normalize_expt(mm38_dlgn_subset, convert = "cpm", norm = "quant", transform = "log2")


pca_mm38_dlgn_subset_norm_nobatch <- plot_pca(mm38_dlgn_subset_norm_nobatch)
pca_norm_plot_nobatch <-  ggplot(pca_mm38_dlgn_subset_norm_nobatch$table, aes(x = PC1, y = PC2, label = labels)) + 
  geom_point(aes(color = condition), 
                      size = 5) +
  scale_color_manual(values = color_values, 
                    aesthetics = "color") + 
  theme_classic() +
  labs(color="Cell Compartment/Genotype")   +    
  ggtitle("PCA with No Batch Correction")  +
  labs(x = paste0("PC1: ",  pca_mm38_dlgn_subset_norm_nobatch$prop_var[1], "% variance"), 
       y =  paste0("PC2: ",  pca_mm38_dlgn_subset_norm_nobatch$prop_var[2], "% variance"))  +
  geom_label_repel(data=filter(pca_mm38_dlgn_subset_norm_nobatch$table, 
                               labels %in% c('S3_Ht_dLGN', 'S4_Ht_dLGN')),
                   nudge_x = 0.1,
                   nudge_y = 0.1)


pca_mm38_dlgn_subset_norm_batch <- plot_pca(mm38_dlgn_subset_norm_batch)
pca_norm_plot_batch <-  ggplot(pca_mm38_dlgn_subset_norm_batch$table, aes(x = PC1, y = PC2, label = labels)) + 
  geom_point(aes(color = condition), 
                      size = 5) +
  scale_color_manual(values = color_values, 
                    aesthetics = "color") + 
  theme_classic() +
  labs(color="Cell Compartment/Genotype")   +    
  ggtitle("PCA with SVA Batch Correction")  +
  labs(x = paste0("PC1: ",  pca_mm38_dlgn_subset_norm_batch$prop_var[1], "% variance"), 
       y =  paste0("PC2: ",  pca_mm38_dlgn_subset_norm_batch$prop_var[2], "% variance"))  +
  geom_label_repel(data=filter(pca_mm38_dlgn_subset_norm_batch$table, 
                               labels %in% c('S3_Ht_dLGN', 'S4_Ht_dLGN')),
                   nudge_x = 0.1,
                   nudge_y = 0.1)


pca_norm_plot_nobatch
pca_norm_plot_batch
```


### Volcano Plots ad DE table results

```{r}
keepers <- names(mm_de_dlgn_sva$deseq$all_tables)

res_tbls_dlgn_sva <- c()
for (tbl in keepers){
  deseq_tbl <- mm_de_dlgn_sva$deseq$all_tables[[tbl]]
  rownames(deseq_tbl) <- paste0("gene:", rownames(deseq_tbl))
  res_tbls_dlgn_sva[[tbl]] <- merge(deseq_tbl, hisat_annot, by = "row.names")
  res_tbls_dlgn_sva[[tbl]] <- set_sig_limma(res_tbls_dlgn_sva[[tbl]], factors = c(paste(gsub("\\_.*","",tbl), "\nEnriched"), 
                                                                paste(sub('.*\\_', '', tbl), "\nEnriched")))
  res_tbls_dlgn_sva[[paste0(tbl, "volc_plotly")]] <- volc_plot(res_tbls_dlgn_sva[[tbl]], 
                                                      title = paste("Volcano Plot:", tbl), 
                                                      type = "plotly", 
                                                      tbl = "limma",
                                                      gene_name_col = "external_gene_name")
  res_tbls_dlgn_sva[[paste0(tbl, "volc_ggplot")]] <- volc_plot(res_tbls_dlgn_sva[[tbl]], 
                                                      title = paste("Volcano Plot:", tbl), 
                                                      tbl = "limma")
}

res_tbls_dlgn_sva$dLGNko_vs_dLGNhetvolc_ggplot
res_tbls_dlgn_sva$dLGNko_vs_dLGNhetvolc_plotly

res_tbls_dlgn_sva$dLGNko_vs_dLGNhet %>%
    filter(adj.P.Val <= 0.05) %>%
    filter(abs(logFC) >= 1) %>%
    select("ensembl_gene_id", "external_gene_name", "Significance", "description", "baseMean", "logFC", "adj.P.Val") %>%
    arrange(logFC) %>%
    DT::datatable(rownames = FALSE)
```

Samd11 and Dmrt3 are still both upregulated in dLGN KOs (as they were without including the high-expressors). 

Pttg1, Sema3g, Ifih1, Dppa5a, and Prg4 are all upregulated in dLGN Hets including and excluding the high expressors. We do lose a few genes when including high expressors that were enriched in dLGN hets before.




## SCN

### PCA pre and post SVAseq

```{r}
mm38_scn_subset_norm_batch <- normalize_expt(mm38_scn_subset, convert = "cpm", transform = "log2", batch = "svaseq")
mm38_scn_subset_norm_nobatch <- normalize_expt(mm38_scn_subset, convert = "cpm", norm = "quant", transform = "log2")


pca_mm38_scn_subset_norm_nobatch <- plot_pca(mm38_scn_subset_norm_nobatch)
pca_norm_plot_nobatch_scn <-  ggplot(pca_mm38_scn_subset_norm_nobatch$table, aes(x = PC1, y = PC2, label = labels)) + 
  geom_point(aes(color = condition), 
                      size = 5) +
  scale_color_manual(values = color_values[5:6], 
                    aesthetics = "color") + 
  theme_classic() +
  labs(color="Cell Compartment/Genotype")   +    
  ggtitle("PCA with No Batch Correction")  +
  labs(x = paste0("PC1: ",  pca_mm38_scn_subset_norm_nobatch$prop_var[1], "% variance"), 
       y =  paste0("PC2: ",  pca_mm38_scn_subset_norm_nobatch$prop_var[2], "% variance"))  +
  geom_label_repel(data=filter(pca_mm38_scn_subset_norm_nobatch$table, 
                               labels %in% c('S3_Het_SCN', 'S4_Het_SCN')),
                   nudge_x = 0.2,
                   nudge_y = 0.2)


pca_mm38_scn_subset_norm_batch <- plot_pca(mm38_scn_subset_norm_batch)
pca_norm_plot_batch_scn <-  ggplot(pca_mm38_scn_subset_norm_batch$table, aes(x = PC1, y = PC2, label = labels)) + 
  geom_point(aes(color = condition), 
                      size = 5) +
  scale_color_manual(values = color_values[5:6], 
                    aesthetics = "color") + 
  theme_classic() +
  labs(color="Cell Compartment/Genotype")   +    
  ggtitle("PCA with SVA Batch Correction")  +
  labs(x = paste0("PC1: ",  pca_mm38_scn_subset_norm_batch$prop_var[1], "% variance"), 
       y =  paste0("PC2: ",  pca_mm38_scn_subset_norm_batch$prop_var[2], "% variance"))  +
  geom_label_repel(data=filter(pca_mm38_scn_subset_norm_batch$table, 
                               labels %in% c('S3_Het_SCN', 'S4_Het_SCN')),
                   nudge_x = 0.2,
                   nudge_y = 0.2)


pca_norm_plot_nobatch_scn
pca_norm_plot_batch_scn
```


### Volcano Plots ad DE table results

```{r}
keepers <- names(mm_de_scn_sva$deseq$all_tables)

res_tbls_scn_sva <- c()
for (tbl in keepers){
  deseq_tbl <- mm_de_scn_sva$deseq$all_tables[[tbl]]
  rownames(deseq_tbl) <- paste0("gene:", rownames(deseq_tbl))
  res_tbls_scn_sva[[tbl]] <- merge(deseq_tbl, hisat_annot, by = "row.names")
  res_tbls_scn_sva[[tbl]] <- set_sig_limma(res_tbls_scn_sva[[tbl]], factors = c(paste(gsub("\\_.*","",tbl), "\nEnriched"), 
                                                                paste(sub('.*\\_', '', tbl), "\nEnriched")))
  res_tbls_scn_sva[[paste0(tbl, "volc_plotly")]] <- volc_plot(res_tbls_scn_sva[[tbl]], 
                                                      title = paste("Volcano Plot:", tbl), 
                                                      type = "plotly", 
                                                      tbl = "limma",
                                                      gene_name_col = "external_gene_name")
  res_tbls_scn_sva[[paste0(tbl, "volc_ggplot")]] <- volc_plot(res_tbls_scn_sva[[tbl]], 
                                                      title = paste("Volcano Plot:", tbl), 
                                                      tbl = "limma")
}

res_tbls_scn_sva$SCNko_vs_SCNhetvolc_ggplot
res_tbls_scn_sva$SCNko_vs_SCNhetvolc_plotly

res_tbls_scn_sva$SCNko_vs_SCNhet %>%
    filter(adj.P.Val <= 0.05) %>%
    filter(abs(logFC) >= 1) %>%
    select("ensembl_gene_id", "external_gene_name", "Significance", "description", "baseMean", "logFC", "adj.P.Val") %>%
    arrange(logFC) %>%
    DT::datatable(rownames = FALSE)
```



## Retina

### PCA pre and post SVAseq

```{r}
mm38_ret_subset_norm_batch <- normalize_expt(mm38_ret_subset, convert = "cpm", transform = "log2", batch = "svaseq")
mm38_ret_subset_norm_nobatch <- normalize_expt(mm38_ret_subset, convert = "cpm", norm = "quant", transform = "log2")


pca_mm38_ret_subset_norm_nobatch <- plot_pca(mm38_ret_subset_norm_nobatch)
pca_norm_plot_nobatch_ret <-  ggplot(pca_mm38_ret_subset_norm_nobatch$table, aes(x = PC1, y = PC2, label = labels)) + 
  geom_point(aes(color = condition), 
                      size = 5) +
  scale_color_manual(values = color_values[3:4], 
                    aesthetics = "color") + 
  theme_classic() +
  labs(color="Cell Compartment/Genotype")   +    
  ggtitle("PCA with No Batch Correction")  +
  labs(x = paste0("PC1: ",  pca_mm38_ret_subset_norm_nobatch$prop_var[1], "% variance"), 
       y =  paste0("PC2: ",  pca_mm38_ret_subset_norm_nobatch$prop_var[2], "% variance"))  +
  geom_label_repel(data=filter(pca_mm38_ret_subset_norm_nobatch$table, 
                               labels %in% c('S3_Het_Ret', 'S4_Het_Ret')),
                   nudge_x = 0.2,
                   nudge_y = 0.2)


pca_mm38_ret_subset_norm_batch <- plot_pca(mm38_ret_subset_norm_batch)
pca_norm_plot_batch_ret <-  ggplot(pca_mm38_ret_subset_norm_batch$table, aes(x = PC1, y = PC2, label = labels)) + 
  geom_point(aes(color = condition), 
                      size = 5) +
  scale_color_manual(values = color_values[3:4], 
                    aesthetics = "color") + 
  theme_classic() +
  labs(color="Cell Compartment/Genotype")   +    
  ggtitle("PCA with SVA Batch Correction")  +
  labs(x = paste0("PC1: ",  pca_mm38_ret_subset_norm_batch$prop_var[1], "% variance"), 
       y =  paste0("PC2: ",  pca_mm38_ret_subset_norm_batch$prop_var[2], "% variance"))  +
  geom_label_repel(data=filter(pca_mm38_ret_subset_norm_batch$table, 
                               labels %in% c('S3_Het_Ret', 'S4_Het_Ret')),
                   nudge_x = 0.2,
                   nudge_y = 0.2)


pca_norm_plot_nobatch_ret
pca_norm_plot_batch_ret
```


### Volcano Plots ad DE table results

```{r}
keepers <- names(mm_de_ret_sva$deseq$all_tables)

res_tbls_ret_sva <- c()
for (tbl in keepers){
  deseq_tbl <- mm_de_ret_sva$deseq$all_tables[[tbl]]
  rownames(deseq_tbl) <- paste0("gene:", rownames(deseq_tbl))
  res_tbls_ret_sva[[tbl]] <- merge(deseq_tbl, hisat_annot, by = "row.names")
  res_tbls_ret_sva[[tbl]] <- set_sig_limma(res_tbls_ret_sva[[tbl]], factors = c(paste(gsub("\\_.*","",tbl), "\nEnriched"), 
                                                                paste(sub('.*\\_', '', tbl), "\nEnriched")))
  res_tbls_ret_sva[[paste0(tbl, "volc_plotly")]] <- volc_plot(res_tbls_ret_sva[[tbl]], 
                                                      title = paste("Volcano Plot:", tbl), 
                                                      type = "plotly", 
                                                      tbl = "limma",
                                                      gene_name_col = "external_gene_name")
  res_tbls_ret_sva[[paste0(tbl, "volc_ggplot")]] <- volc_plot(res_tbls_ret_sva[[tbl]], 
                                                      title = paste("Volcano Plot:", tbl), 
                                                      tbl = "limma")
}

res_tbls_ret_sva$Retinako_vs_Retinahetvolc_ggplot
res_tbls_ret_sva$Retinako_vs_Retinahetvolc_plotly

res_tbls_ret_sva$Retinako_vs_Retinahet %>%
    filter(adj.P.Val <= 0.05) %>%
    filter(abs(logFC) >= 1) %>%
    select("ensembl_gene_id", "external_gene_name", "Significance", "description", "baseMean", "logFC", "adj.P.Val") %>%
    arrange(logFC) %>%
    DT::datatable(rownames = FALSE)
```














